---
title: "Dataset Reference Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dataset Reference Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(psyc350data)
```

This guide provides a quick reference for every dataset in the psyc350data package, including variable descriptions, categorical coding schemes, available export functions, and suggested analyses.

---

## Superman Actor Data

**Dataset:** `superman` | **Export:** `export_superman_sav()`
**Rows:** 11 | **Variables:** 21

Data on actors who have portrayed Clark Kent/Superman across film, TV, and serial media.
```{r}
str(superman)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `media` | 1 | Superman 2025 |
| `media` | 2 | Superman: The Movie |
| `media` | 3 | Smallville |
| `media` | 4 | Superman Returns |
| `media` | 5 | Superman & the Mole Men |
| `media` | 6 | Man of Steel |
| `media` | 7 | Superman |
| `media` | 8 | Superman & Lois |
| `media` | 9 | Lois & Clark |
| `media` | 10 | The Adventures of Superboy |
| `type` | 1 | Film |
| `type` | 2 | TV Series |
| `type` | 3 | Serial |
| `clark_grp` | 1 | Under 6ft (<72 inches) |
| `clark_grp` | 2 | Over 6ft (≥72 inches) |
| `tomatometer` | 1 | Rotten |
| `tomatometer` | 2 | Fresh |
| `height_gap` | 1 | Minimal (<6 inches) |
| `height_gap` | 2 | Average (6-8 inches) |
| `height_gap` | 3 | Big (>8 inches) |
| `popular` | 1 | Low (<1,000 likes) |
| `popular` | 2 | Mid (1,000-100,000 likes) |
| `popular` | 3 | High (>100,000 likes) |

### Export to SPSS
```{r eval = FALSE}
export_superman_sav("superman_data.sav")
# or: export_sav("superman", path = "superman_data.sav")
```

### Suggested Analyses
- Descriptive statistics for height and rating variables
- Correlation between critic and audience scores
- Group comparisons by media type

---

## Superman SMES Data

**Dataset:** `superman_smes` | **Export:** `export_superman_smes_sav()`
**Rows:** 47 | **Variables:** 5

Simulated participant ratings on the Subjective Media Experience Scale (SMES), grouped by the height gap between Superman and Lois Lane actors.
```{r}
str(superman_smes)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `height_gap` | 1 | Minimal (<6 inches) |
| `height_gap` | 2 | Average (6-8 inches) |
| `height_gap` | 3 | Big (>8 inches) |

### Scale Information

| Subscale | Scoring | Range |
|----------|---------|-------|
| `emotional_impact` | Sum of 4 items (1-5 scale) | 4-20 |
| `aesthetic_appeal` | Sum of 3 items (1-5 scale) | 3-15 |
| `cognitive_engagement` | Mean of 4 items (0-7 scale) | 0-7 |

### Export to SPSS
```{r eval = FALSE}
export_superman_smes_sav("superman_smes_data.sav")
```

### Suggested Analyses
- One-way ANOVA comparing SMES subscales across height gap groups
- Post-hoc pairwise comparisons

---

## Hot Ones Guest Data

**Dataset:** `hot_ones` | **Export:** `export_hot_ones_sav()`
**Variables:** 23

Data on guests from the YouTube show Hot Ones, including Scoville ratings for each hot sauce and YouTube engagement metrics.
```{r}
str(hot_ones)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `gender` | 1 | Male |
| `gender` | 2 | Female |
| `result` | 1 | Succeeded |
| `result` | 2 | Failed |
| `occupation` | 1 | Rapper |
| `occupation` | 2 | Athlete |
| `occupation` | 3 | Actor |
| `occupation` | 4 | Actor-Comedian |
| `occupation` | 5 | Comedian |
| `occupation` | 6 | Chef |
| `occupation` | 7 | Actor-Musician |
| `occupation` | 8 | Musician |
| `occupation` | 9 | DJ |
| `occupation` | 10 | YouTuber |
| `occupation` | 11 | Model |
| `occupation` | 12 | Wrestler |
| `occupation` | 13 | Magician |
| `occupation` | 14 | Other |

### Export to SPSS
```{r eval = FALSE}
export_hot_ones_sav("hot_ones_data.sav")
```

### Suggested Analyses
- Chi-square test of success rate by gender or occupation
- Correlation between YouTube views and season number
- Repeated measures across sauce Scoville levels

---

## Tip-Jokes Experiment Data

**Dataset:** `tip_jokes` | **Export:** `export_tip_jokes_sav()`

Experimental data from Gueguen (2002) examining whether a waiter leaving a joke or advertisement on a card affects tipping.
```{r}
str(tip_jokes)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `card` | 1 | Advertisement card |
| `card` | 2 | Joke card |
| `card` | 3 | No card |
| `tip` | 0 | No tip |
| `tip` | 1 | Customer left a tip |
| `ad` | 0 | No ad card |
| `ad` | 1 | Ad card left |
| `joke` | 0 | No joke card |
| `joke` | 1 | Joke card left |
| `none` | 0 | Ad or joke card left |
| `none` | 1 | No card left |

### Export to SPSS
```{r eval = FALSE}
export_tip_jokes_sav("tip_jokes_data.sav")
```

### Suggested Analyses
- Chi-square test of independence (card type × tipping)
- Logistic regression predicting tip from card type

---

## MCU Films Data

**Dataset:** `mcu` | **Export:** `export_mcu_sav()`

Marvel Cinematic Universe films through the Infinity Saga with box office and Rotten Tomatoes data.
```{r}
str(mcu)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `phase` | 1 | Phase 1 |
| `phase` | 2 | Phase 2 |
| `phase` | 3 | Phase 3 |
| `favor` | 1 | Critics score higher |
| `favor` | 2 | Audience score higher |

### Export to SPSS
```{r eval = FALSE}
export_mcu_sav("mcu_data.sav")
```

### Suggested Analyses
- One-way ANOVA comparing box office by phase
- Paired comparison of critics vs. audience scores
- Regression predicting worldwide gross

---

## Mock Jury Data

**Dataset:** `mock_jury` | **Export:** `export_mock_jury_sav()`

Data from Plaster (1989) on effects of defendant attractiveness on mock jury sentencing decisions.
```{r}
str(mock_jury)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `attr` | 1 | Beautiful |
| `attr` | 2 | Average |
| `attr` | 3 | Unattractive |
| `crime` | 1 | Burglary |
| `crime` | 2 | Swindle |

### Export to SPSS
```{r eval = FALSE}
export_mock_jury_sav("mock_jury_data.sav")
```

### Suggested Analyses
- Two-way ANOVA (attractiveness × crime type) on sentence length
- Correlation among photo rating variables
- Multiple regression predicting sentence from photo ratings

---

## Candy Rankings Data

**Full dataset:** `candy` (85 rows × 13 variables) | **Export:** `export_candy_sav()`
**Simplified:** `candy_simple` (85 rows × 5 variables) | **Export:** `export_candy_simple_sav()`

Candy power rankings based on 269,000 head-to-head matchups.
```{r}
str(candy_simple)
```

### Binary Variable Codings (full dataset)

All ingredient variables: 0 = No, 1 = Yes

Applies to: `chocolate`, `fruity`, `caramel`, `peanutyalmondy`, `nougat`, `crispedricewafer`, `hard`, `bar`, `pluribus`

### Export to SPSS
```{r eval = FALSE}
# Full dataset with all ingredient variables
export_candy_sav("candy_data.sav")

# Simplified dataset (5 variables only)
export_candy_simple_sav("candy_simple_data.sav")
```

### Suggested Analyses
- Independent samples t-test (chocolate vs. non-chocolate win %)
- Multiple regression predicting win percentage
- Correlation between sugar, price, and win percentages

---


## Football Concussion Data

**Dataset:** `football` | **Export:** `export_football_sav()`

Brain measurements comparing football players to controls, from Singh et al. (2014).
```{r}
str(football)
```

### Categorical Codings

| Variable | Code | Label |
|----------|------|-------|
| `group` | 1 | Control (no football) |
| `group` | 2 | Football, no concussions |
| `group` | 3 | Football with concussion history |

### Export to SPSS
```{r eval = FALSE}
export_football_sav("football_data.sav")
```

