---
title: "Dataset Reference Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dataset Reference Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(psych350data)
```

This guide provides a quick reference for every dataset in the psych350data package, including variable descriptions, categorical coding schemes, available export functions, and suggested analyses.

---

## Superman Actor Data

**Dataset:** `superman` | **Export:** `export_superman_sav()`
**Rows:** 11 | **Variables:** 21

Data on actors who have portrayed Clark Kent/Superman across film, TV, and serial media.
```{r}
str(superman)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `media` | 1 = Superman 2025, 2 = Superman: The Movie, 3 = Smallville, 4 = Superman Returns, 5 = Superman & the Mole Men, 6 = Man of Steel, 7 = Superman, 8 = Superman & Lois, 9 = Lois & Clark, 10 = The Adventures of Superboy |
| `type` | 1 = Film, 2 = TV Series, 3 = Serial |
| `clark_grp` | 1 = Under 6ft (<72 inches), 2 = Over 6ft (≥72 inches) |
| `tomatometer` | 1 = Rotten, 2 = Fresh |
| `height_gap` | 1 = Minimal (<6 inches), 2 = Average (6-8 inches), 3 = Big (>8 inches) |
| `popular` | 1 = Low (<1,000 likes), 2 = Mid (1,000-100,000 likes), 3 = High (>100,000 likes) |

### Export to SPSS
```{r eval = FALSE}
export_superman_sav("superman_data.sav")
# or: export_sav("superman", path = "superman_data.sav")
```

### Suggested Analyses
- Descriptive statistics for height and rating variables
- Correlation between critic and audience scores
- Group comparisons by media type

---

## Superman SMES Data

**Dataset:** `superman_smes` | **Export:** `export_superman_smes_sav()`
**Rows:** 47 | **Variables:** 5

Simulated participant ratings on the Subjective Media Experience Scale (SMES), grouped by the height gap between Superman and Lois Lane actors.
```{r}
str(superman_smes)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `height_gap` | 1 = Minimal (<6 inches), 2 = Average (6-8 inches), 3 = Big (>8 inches) |

### Scale Information

| Subscale | Scoring | Range |
|----------|---------|-------|
| `emotional_impact` | Sum of 4 items (1-5 scale) | 4-20 |
| `aesthetic_appeal` | Sum of 3 items (1-5 scale) | 3-15 |
| `cognitive_engagement` | Mean of 4 items (0-7 scale) | 0-7 |

### Export to SPSS
```{r eval = FALSE}
export_superman_smes_sav("superman_smes_data.sav")
```

### Suggested Analyses
- One-way ANOVA comparing SMES subscales across height gap groups
- Post-hoc pairwise comparisons

---

## Hot Ones Guest Data

**Dataset:** `hot_ones` | **Export:** `export_hot_ones_sav()`
**Variables:** 23

Data on guests from the YouTube show Hot Ones, including Scoville ratings for each hot sauce and YouTube engagement metrics.
```{r}
str(hot_ones)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `gender` | 1 = Male, 2 = Female |
| `result` | 1 = Succeeded, 2 = Failed |
| `occupation` | 1 = Rapper, 2 = Athlete, 3 = Actor, 4 = Actor-Comedian, 5 = Comedian, 6 = Chef, 7 = Actor-Musician, 8 = Musician, 9 = DJ, 10 = YouTuber, 11 = Model, 12 = Wrestler, 13 = Magician, 14 = Other |

### Export to SPSS
```{r eval = FALSE}
export_hot_ones_sav("hot_ones_data.sav")
```

### Suggested Analyses
- Chi-square test of success rate by gender or occupation
- Correlation between YouTube views and season number
- Repeated measures across sauce Scoville levels

---

## Tip-Jokes Experiment Data

**Dataset:** `tip_jokes` | **Export:** `export_tip_jokes_sav()`

Experimental data from Gueguen (2002) examining whether a waiter leaving a joke or advertisement on a card affects tipping.
```{r}
str(tip_jokes)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `card` | 1 = Advertisement card, 2 = Joke card, 3 = No card |
| `tip` | 0 = No tip, 1 = Customer left a tip |
| `ad` | 0 = No ad card, 1 = Ad card left |
| `joke` | 0 = No joke card, 1 = Joke card left |
| `none` | 0 = Ad or joke card left, 1 = No card left |

### Export to SPSS
```{r eval = FALSE}
export_tip_jokes_sav("tip_jokes_data.sav")
```

### Suggested Analyses
- Chi-square test of independence (card type × tipping)
- Logistic regression predicting tip from card type

---

## MCU Films Data

**Dataset:** `mcu` | **Export:** `export_mcu_sav()`

Marvel Cinematic Universe films through the Infinity Saga with box office and Rotten Tomatoes data.
```{r}
str(mcu)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `phase` | 1 = Phase 1, 2 = Phase 2, 3 = Phase 3 |
| `favor` | 1 = Critics score higher, 2 = Audience score higher |

### Export to SPSS
```{r eval = FALSE}
export_mcu_sav("mcu_data.sav")
```

### Suggested Analyses
- One-way ANOVA comparing box office by phase
- Paired comparison of critics vs. audience scores
- Regression predicting worldwide gross

---

## Mock Jury Data

**Dataset:** `mock_jury` | **Export:** `export_mock_jury_sav()`

Data from Plaster (1989) on effects of defendant attractiveness on mock jury sentencing decisions.
```{r}
str(mock_jury)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `attr` | 1 = Beautiful, 2 = Average, 3 = Unattractive |
| `crime` | 1 = Burglary, 2 = Swindle |

### Export to SPSS
```{r eval = FALSE}
export_mock_jury_sav("mock_jury_data.sav")
```

### Uses: 
- Two-way ANOVA (attractiveness × crime type) on sentence length
- Correlation among photo rating variables
- Multiple regression predicting sentence from photo ratings

---

## Candy Rankings Data

**Full dataset:** `candy` (85 rows × 13 variables) | **Export:** `export_candy_sav()`
**Simplified:** `candy_simple` (85 rows × 5 variables) | **Export:** `export_candy_simple_sav()`

Candy power rankings based on 269,000 head-to-head matchups.
```{r}
str(candy_simple)
```

### Binary Variable Codings (full dataset)

| Variable | Labels |
|----------|--------|
| `chocolate` | 0 = No, 1 = Yes |
| `fruity` | 0 = No, 1 = Yes |
| `caramel` | 0 = No, 1 = Yes |
| `peanutyalmondy` | 0 = No, 1 = Yes |
| `nougat` | 0 = No, 1 = Yes |
| `crispedricewafer` | 0 = No, 1 = Yes |
| `hard` | 0 = No, 1 = Yes |
| `bar` | 0 = No, 1 = Yes |
| `pluribus` | 0 = No, 1 = Yes |

### Export to SPSS
```{r eval = FALSE}
# Full dataset with all ingredient variables
export_candy_sav("candy_data.sav")

# Simplified dataset (5 variables only)
export_candy_simple_sav("candy_simple_data.sav")
```

### Suggested Analyses
- Independent samples t-test (chocolate vs. non-chocolate win %)
- Multiple regression predicting win percentage
- Correlation between sugar, price, and win percentages

---

## Football Concussion Data

**Dataset:** `football` | **Export:** `export_football_sav()`

Brain measurements comparing football players to controls, from Singh et al. (2014).
```{r}
str(football)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `group` | 1 = Control (no football), 2 = Football no concussions, 3 = Football with concussion history |

### Export to SPSS
```{r eval = FALSE}
export_football_sav("football_data.sav")
```

### Suggested Analyses
- One-way ANOVA comparing hippocampus volume across groups
- Post-hoc pairwise comparisons
- Correlation between years played and brain volume

---

## Nebraska Football Box Scores

**Dataset:** `huskers` | **Export:** `export_huskers_sav()`

Game-level team statistics for all Nebraska Cornhuskers football games from September 1962 through the 2024 season, including scoring, rushing, passing, turnovers, penalties, point spreads, and weather data.
```{r}
str(huskers)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `result` | W = Win, L = Loss, T = Tie |
| `site` | home = Home, away = Away, neutral-home = Neutral (home), neutral-away = Neutral (away) |
| `conference` | 0 = Non-conference, 1 = Conference game |

### Data Completeness Notes

- Scoring by quarter, first down, third/fourth down, and opponent penalty data begins in 2004
- Nebraska penalty data incomplete before 1972-09-16
- Time of possession data begins in 2012, complete from 2013 onward
- Point spread data mostly absent before 1978
- Betting total data available for most games from 2006 onward

### Export to SPSS
```{r eval = FALSE}
export_huskers_sav("huskers_data.sav")
```

### Suggested Analyses
- Regression predicting Nebraska score from rushing/passing yards
- Comparing home vs away performance
- Time series analysis of performance across seasons
- Weather effects on scoring

---

## Interpersonal Relationships Survey Data

**Dataset:** `interpersonal_data` | **Export:** `export_interpersonal_sav()`
**Rows:** 574 | **Variables:** 33

Simulated survey data from college students containing demographics, relationship variables, and scores on several interpersonal and psychological scales including the DAQ, Interpersonal Dependency Inventory, Interpersonal Reactivity Index, Sociability Scale, RISC, GCB, and LSAS-SR.
```{r}
str(interpersonal_data)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `gender` | 1 = Male, 2 = Female, 3 = Another |
| `sexorient` | 1 = Asexual, 2 = Bisexual, 3 = Demisexual, 4 = Gay, 5 = Lesbian, 6 = Pansexual, 7 = Queer, 8 = Questioning, 9 = Sexually fluid, 10 = Straight or heterosexual, 11 = Other |
| `race` | 1 = Asian, 2 = Black, 3 = Indigenous/Aboriginal/First Nations, 4 = Latino/Hispanic, 5 = Middle Eastern, 6 = White, 7 = Other |
| `hand` | 1 = Right, 2 = Left, 3 = Both |
| `community` | 1 = Rural, 2 = Small town, 3 = Suburban, 4 = Urban |
| `parentedu` | 0 = No, 1 = Yes |
| `famclass` | 1 = Working class, 2 = Lower class, 3 = Lower middle class, 4 = Upper middle class, 5 = Upper class |
| `greek_in` | 1 = Independent, 2 = Greek |
| `campus` | 0 = No, 1 = Yes |
| `relsp` | 1 = In a monogamous relationship, 2 = In a polyamorous relationship, 3 = In multiple relationships, 4 = Not in a relationship, 5 = I prefer not to answer |

### Export to SPSS
```{r eval = FALSE}
export_interpersonal_sav("interpersonal_data.sav")
```

### Uses: 
- Correlation matrix of interpersonal scales
- Regression predicting social anxiety from dependency measures
- Group comparisons by relationship status

---

## Self-Descriptive Survey Data

**Dataset:** `self_descriptive_data` | **Export:** `export_selfdescriptive_sav()`
**Rows:** 547 | **Variables:** 37

Simulated survey data from college students containing demographics, relationship variables, and scores on personality and self-concept measures including the TIPI Big Five, MAAS, RFQ, ATQ-30, Rosenberg Self-Esteem, and New General Self-Efficacy scales.
```{r}
str(self_descriptive_data)
```

### Categorical Codings

| Variable | Labels |
|----------|--------|
| `gender` | 1 = Male, 2 = Female, 3 = Another |
| `sexorient` | 1 = Asexual, 2 = Bisexual, 3 = Demisexual, 4 = Gay, 5 = Lesbian, 6 = Pansexual, 7 = Queer, 8 = Questioning, 9 = Sexually fluid, 10 = Straight or heterosexual, 11 = Other |
| `race` | 1 = Asian, 2 = Black, 3 = Indigenous/Aboriginal/First Nations, 4 = Latino/Hispanic, 5 = Middle Eastern, 6 = White, 7 = Other |
| `hand` | 1 = Right, 2 = Left, 3 = Both |
| `community` | 1 = Rural, 2 = Small town, 3 = Suburban, 4 = Urban |
| `parentedu` | 0 = No, 1 = Yes |
| `famclass` | 1 = Working class, 2 = Lower class, 3 = Lower middle class, 4 = Upper middle class, 5 = Upper class |
| `greek_in` | 1 = Independent, 2 = Greek |
| `campus` | 0 = No, 1 = Yes |
| `relsp` | 1 = In a monogamous relationship, 2 = In a polyamorous relationship, 3 = In multiple relationships, 4 = Not in a relationship, 5 = I prefer not to answer |

### Export to SPSS
```{r eval = FALSE}
export_selfdescriptive_sav("selfdescriptive_data.sav")
```
