---
title: "Dataset Reference Guide"
---
```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(psych350data)
library(dplyr)
```

This guide provides a quick reference for every dataset in the psych350data package.

## How Categorical Variables Work

**In R:** Categorical variables use human-readable character values (e.g., `"Film"`, `"TV Series"`, `"Minimal"`).

**In SPSS exports:** These are automatically converted to numeric codes with value labels.
```{r}
# Example: categorical variables are human-readable in R
superman |>
  select(num, media, type, height_gap) |>
  head(4)
```

---

## Superman Actor Data

**Dataset:** `superman` | **Export:** `export_superman_sav()`

Data on actors who have portrayed Clark Kent/Superman across film, TV, and serial media.
```{r}
glimpse(superman)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `type` | 1 = Film, 2 = TV Series, 3 = Serial |
| `clark_grp` | 1 = Under 6ft, 2 = 6ft or taller |
| `height_gap` | 1 = Minimal, 2 = Average, 3 = Big |
| `tomatometer` | 1 = Rotten, 2 = Fresh |
| `popular` | 1 = Low, 2 = Mid, 3 = High |

---

## Superman SMES Data

**Dataset:** `superman_smes` | **Export:** `export_superman_smes_sav()`

Simulated participant ratings on the Subjective Media Experience Scale (SMES).
```{r}
glimpse(superman_smes)
```

### Scale Information

| Subscale | Range |
|----------|-------|
| `emotional_impact` | 4-20 |
| `aesthetic_appeal` | 3-15 |
| `cognitive_engagement` | 0-7 |

---

## Superman Movies Data

**Dataset:** `superman_movies` | **Export:** `export_superman_movies_sav()`

Box office and production data for Superman theatrical films.
```{r}
glimpse(superman_movies)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `mpaa` | 1 = G, 2 = PG, 3 = PG-13, 4 = R |
| `budget_cat` | 1 = Low, 2 = Medium, 3 = High |
| `box_office_cat` | 1 = Low, 2 = Medium, 3 = High |

---

## Hot Ones Guest Data

**Dataset:** `hot_ones` | **Export:** `export_hot_ones_sav()`

Data on guests from the YouTube show Hot Ones.
```{r}
glimpse(hot_ones)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `gender` | 1 = Male, 2 = Female |
| `result` | 1 = Succeeded, 2 = Failed, 3 = Incomplete |
| `occupation` | 1-14 (Rapper, Athlete, Actor, etc.) |

---

## Hot Ones Sauces Data

**Dataset:** `hot_ones_sauces` | **Export:** `export_hot_ones_sauces_sav()`

Hot sauce data by season and position in the lineup.
```{r}
glimpse(hot_ones_sauces)
```

---

## Hot Ones Episodes Data

**Dataset:** `hot_ones_episodes` | **Export:** `export_hot_ones_episodes_sav()`

Episode-level data with YouTube engagement metrics.
```{r}
glimpse(hot_ones_episodes)
```

---

## Tip-Jokes Experiment Data

**Dataset:** `tip_jokes` | **Export:** `export_tip_jokes_sav()`

Experimental data from Gueguen (2002) examining whether a waiter leaving a joke or advertisement on a card affects tipping.
```{r}
glimpse(tip_jokes)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `card` | 1 = Advertisement, 2 = Joke, 3 = None |
| `tip` | 0 = No, 1 = Yes |

---

## MCU Films Data

**Dataset:** `mcu` | **Export:** `export_mcu_sav()`

Marvel Cinematic Universe films through the Infinity Saga.
```{r}
glimpse(mcu)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `phase` | 1 = Phase 1, 2 = Phase 2, 3 = Phase 3 |
| `favor` | 1 = Critics, 2 = Audience |

---

## Mock Jury Data

**Dataset:** `mock_jury` | **Export:** `export_mock_jury_sav()`

Data from Plaster (1989) on effects of defendant attractiveness on mock jury sentencing.
```{r}
glimpse(mock_jury)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `attr` | 1 = Beautiful, 2 = Average, 3 = Unattractive |
| `crime` | 1 = Burglary, 2 = Swindle |

---

## Candy Rankings Data

**Full dataset:** `candy` | **Export:** `export_candy_sav()`
**Simplified:** `candy_simple` | **Export:** `export_candy_simple_sav()`

Candy power rankings based on 269,000 head-to-head matchups.
```{r}
glimpse(candy_simple)
```

### Binary Variables (SPSS codes)

All binary variables: 0 = No, 1 = Yes

Variables: `chocolate`, `fruity`, `caramel`, `peanutyalmondy`, `nougat`, `crispedricewafer`, `hard`, `bar`, `pluribus`

---

## Football Concussion Data

**Dataset:** `football` | **Export:** `export_football_sav()`

Brain measurements comparing football players to controls (Singh et al., 2014).
```{r}
glimpse(football)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `group` | 1 = Control, 2 = Football no concussion, 3 = Football with concussion |

---

## Nebraska Football Box Scores

**Dataset:** `huskers` | **Export:** `export_huskers_sav()`

Game-level statistics for Nebraska Cornhuskers football (1962-2024).
```{r}
glimpse(huskers)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `result` | 1 = Win, 2 = Loss, 3 = Tie |
| `site` | 1 = Home, 2 = Away, 3 = Neutral (home), 4 = Neutral (away) |
| `conference` | 0 = Non-conference, 1 = Conference |

---

## Interpersonal Relationships Survey Data

**Dataset:** `interpersonal_data` | **Export:** `export_interpersonal_sav()`

Simulated survey data (n = 574) with demographics and interpersonal scales.
```{r}
glimpse(interpersonal_data)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------|--------|
| `gender` | 1 = Male, 2 = Female, 3 = Another |
| `race` | 1 = Asian, 2 = Black, 3 = Indigenous, 4 = Latino/Hispanic, 5 = Middle Eastern, 6 = White, 7 = Other |
| `hand` | 1 = Right, 2 = Left, 3 = Both |
| `community` | 1 = Rural, 2 = Small town, 3 = Suburban, 4 = Urban |
| `famclass` | 1 = Working, 2 = Lower, 3 = Lower middle, 4 = Upper middle, 5 = Upper |
| `greek_in` | 1 = Independent, 2 = Greek |

### Scale Variables

DAQ subscales, IDI subscales, IRI subscales, Sociability subscales, RISC, GCB, LSAS-SR

---

## Self-Descriptive Survey Data

**Dataset:** `self_descriptive_data` | **Export:** `export_selfdescriptive_sav()`

Simulated survey data (n = 547) with demographics and personality scales.
```{r}
glimpse(self_descriptive_data)
```

### Categorical Variables

Same coding as `interpersonal_data` for demographic variables.

### Scale Variables

TIPI Big Five, MAAS subscales, RFQ subscales, ATQ-30 subscales, Rosenberg Self-Esteem, NGSE

---

## Quick Reference

| Dataset | Export Function |
|---------|-----------------|
| `superman` | `export_superman_sav()` |
| `superman_smes` | `export_superman_smes_sav()` |
| `superman_movies` | `export_superman_movies_sav()` |
| `hot_ones` | `export_hot_ones_sav()` |
| `hot_ones_sauces` | `export_hot_ones_sauces_sav()` |
| `hot_ones_episodes` | `export_hot_ones_episodes_sav()` |
| `tip_jokes` | `export_tip_jokes_sav()` |
| `mcu` | `export_mcu_sav()` |
| `mock_jury` | `export_mock_jury_sav()` |
| `candy` | `export_candy_sav()` |
| `candy_simple` | `export_candy_simple_sav()` |
| `football` | `export_football_sav()` |
| `huskers` | `export_huskers_sav()` |
| `interpersonal_data` | `export_interpersonal_sav()` |
| `self_descriptive_data` | `export_selfdescriptive_sav()` |

Use `list_datasets()` to see all available datasets.
