---
title: "Exporting Datasets to SPSS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exporting Datasets to SPSS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

All psyc350data datasets can be exported as SPSS `.sav` files with full metadata — variable labels, value labels, and user-defined missing values. This is useful when you need to work in SPSS for lab assignments.

## Setup
```{r setup}
library(psyc350data)
```

## Exporting a Single Dataset

Use `export_sav()` to save any dataset as a `.sav` file:
```{r eval = FALSE}
# Save to your desktop
export_sav("superman", path = "~/Desktop/superman_data.sav")

# Save to the current working directory (default)
export_sav("mock_jury")
# Creates: mock_jury_data.sav
```

## Exporting All Datasets

Use `export_all_sav()` to export every dataset at once:
```{r eval = FALSE}
# Create a folder with all .sav files
export_all_sav(dir = "~/Desktop/PSYC350_SPSS/")
```

This creates one `.sav` file per dataset:
```
PSYC350_SPSS/
├── superman_data.sav
├── superman_smes_data.sav
├── hot_ones_data.sav
├── tip_jokes_data.sav
├── mcu_data.sav
├── mock_jury_data.sav
├── candy_data.sav
├── candy_simple_data.sav
├── affairs_data.sav
└── football_data.sav
```

## What's in the .sav Files?

The exported files include SPSS metadata that you won't see in the R tibbles:

### Variable Labels

Every variable gets a descriptive label visible in SPSS Variable View. For example, in the superman dataset:

| Variable | Label |
|----------|-------|
| `clark_height` | Height of Clark Kent/Superman actor in meters |
| `rt_critics_score` | Rotten Tomatoes critics score (0-100 scale) |
| `tomatometer` | Whether the media was liked by more than 60% of critics |

### Value Labels

Categorical variables include labeled values. For example:

| Variable | Code | Label |
|----------|------|-------|
| `type` | 1 | Film |
| `type` | 2 | TV Series |
| `type` | 3 | Serial |

### Missing Values

In the R objects, missing data is represented as `NA`. In the `.sav` files, missing values are coded as `-99` and registered as user-defined missing values so SPSS excludes them from analyses automatically.

## Controlling Missing Value Behavior

By default, `export_sav()` converts `NA` to `-99` and sets the SPSS missing value attribute. To write standard SPSS system-missing values instead:
```{r eval = FALSE}
export_sav("superman", path = "superman_sysmis.sav", use_sentinel = FALSE)
```

## Available Datasets

To see all datasets you can export:
```{r}
list_datasets()
```

## Tips for SPSS Users

1. **After opening in SPSS**, check Variable View to see all labels
2. **Missing values** are already defined — no need to set them manually
3. **Categorical variables** will show value labels in Data View if you toggle "Value Labels" in the toolbar
4. If you need to re-export after updating the package, just run `export_sav()` again
```

---

## Update `DESCRIPTION` for Vignettes

Add `knitr` and `rmarkdown` to Suggests:
```
Suggests:
    dplyr,
    readxl,
    tidyr,
    purrr,
    lubridate,
    knitr,
    rmarkdown,
    testthat (>= 3.0.0)
VignetteBuilder: knitr
