---
title: "Exporting Datasets to SPSS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exporting Datasets to SPSS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

All psych350data datasets can be exported as SPSS `.sav` files with full metadata. The export process automatically:

- **Converts categorical variables** from human-readable strings to numeric codes
- **Adds value labels** so SPSS displays category names
- **Replaces `NA` with `-99`** and registers it as user-defined missing
- **Adds variable labels** describing each column

## Setup
```{r setup}
library(psych350data)
library(dplyr)
```

## Quick Export

Each dataset has a dedicated export function:
```{r eval = FALSE}
# Export to current working directory
export_superman_sav()

# Export to specific location
export_superman_sav("~/Desktop/superman_data.sav")
```

Or use the generic function:
```{r eval = FALSE}
export_sav("superman", path = "superman_data.sav")
```

## All Export Functions

| Function | Default filename |
|----------|-----------------|
| `export_superman_sav()` | `superman_data.sav` |
| `export_superman_smes_sav()` | `superman_smes_data.sav` |
| `export_superman_movies_sav()` | `superman_movies_data.sav` |
| `export_superman_combined_sav()` | `superman_combined_data.sav` |
| `export_hot_ones_sav()` | `hot_ones_data.sav` |
| `export_hot_ones_sauces_sav()` | `hot_ones_sauces_data.sav` |
| `export_hot_ones_episodes_sav()` | `hot_ones_episodes_data.sav` |
| `export_tip_jokes_sav()` | `tip_jokes_data.sav` |
| `export_mcu_sav()` | `mcu_data.sav` |
| `export_mock_jury_sav()` | `mock_jury_data.sav` |
| `export_candy_sav()` | `candy_data.sav` |
| `export_candy_simple_sav()` | `candy_simple_data.sav` |
| `export_football_sav()` | `football_data.sav` |
| `export_huskers_sav()` | `huskers_data.sav` |
| `export_interpersonal_sav()` | `interpersonal_data.sav` |
| `export_selfdescriptive_sav()` | `selfdescriptive_data.sav` |

## Exporting a Subset of Variables

Use `dplyr::select()` to choose specific columns before exporting:
```{r eval = FALSE}
# Export selected variables
superman |>
  select(num, media, year, type, clark_height, rt_critics_score) |>
  export_sav(path = "superman_subset.sav")

# Use tidyselect helpers
interpersonal_data |>
  select(age, gender, race, gcb, risc, lsas) |>
  export_sav(path = "interpersonal_subset.sav")

# Select by pattern
huskers |>
  select(date, season, opp, result, starts_with("ne_")) |>
  export_sav(path = "huskers_nebraska_stats.sav")
```

Labels are preserved for all selected variables.

## Export All Datasets at Once
```{r eval = FALSE}
export_all_sav(dir = "~/Desktop/PSYC350_SPSS/")
```

This creates one `.sav` file per dataset in the specified folder.

## What Gets Converted During Export

### Categorical Variables

In R, categorical variables are stored as readable strings:
```{r}
superman |>
  select(num, media, type, height_gap) |>
  head(3)
```

In the exported SPSS file, these become numeric codes with value labels:

| Variable | R Value | SPSS Code | SPSS Label |
|----------|---------|-----------|------------|
| `type` | "Film" | 1 | Film |
| `type` | "TV Series" | 2 | TV Series |
| `type` | "Serial" | 3 | Serial |
| `height_gap` | "Minimal" | 1 | Minimal (<6 inches) |
| `height_gap` | "Average" | 2 | Average (6-8 inches) |
| `height_gap` | "Big" | 3 | Big (>8 inches) |

### Missing Values

In R, missing values are `NA`:
```{r}
superman |>
  select(num, media, ldb_scores) |>
  filter(is.na(ldb_scores))
```

In SPSS, these become `-99` with the missing value attribute set, so SPSS automatically excludes them from analyses.

### Variable Labels

Every variable gets a descriptive label visible in SPSS Variable View:

| Variable | SPSS Variable Label |
|----------|---------------------|
| `clark_height` | Height of Clark Kent/Superman actor in meters |
| `rt_critics_score` | Rotten Tomatoes critics score (0-100 scale) |
| `height_gap` | Relative size of height gap between actors |

## Controlling Missing Value Behavior

By default, `NA` values are converted to `-99` with SPSS missing value codes. To use SPSS system-missing instead:
```{r eval = FALSE}
export_sav("superman", path = "superman_sysmis.sav", use_sentinel = FALSE)
export_superman_sav("superman_sysmis.sav", use_sentinel = FALSE)
```

## Tips for SPSS Users

1. **Check Variable View** after opening to see all labels
2. **Toggle "Value Labels"** in the toolbar to see category names in Data View
3. **Missing values are pre-configured** — no manual setup needed
4. **Re-export anytime** by running the export function again after updating the package
5. **Subset before exporting** if you only need certain variables for an assignment
