---
title: "Getting Started with psych350data"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Getting Started with psych350data}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>"
)
```

## Overview

The psych350data package provides all the datasets you need for PSYC 350 labs. Each dataset is pre-cleaned and ready to use in two ways:

1. **In R** — Datasets use human-readable values (like "Film", "TV Series") and standard `NA` for missing data. Perfect for analysis with psych350lab functions.

2. **In SPSS** — Export functions create `.sav` files with numeric codes, value labels, and `-99` for missing values.

## Loading the Package
```{r setup}
library(psych350data)
library(dplyr)
```

## Browsing Available Datasets

To see every dataset in the package:
```{r}
list_datasets()
```

For detailed documentation on any dataset, use the help system:
```{r eval = FALSE}
?superman
?hot_ones
?mock_jury
```

## Using Datasets in R

Every dataset is a tibble that's immediately available after loading the package. Categorical variables use descriptive character values that are easy to read and work with:
```{r}
# Categorical variables are human-readable
superman |>
 select(num, media, type, clark_grp, height_gap) |>
 head()
```

### Missing Values in R

Missing values are represented as `NA`, which R functions handle automatically:
```{r}
# See which Superman entries have missing Letterboxd scores
superman |>
 select(num, media, ldb_likes, ldb_scores) |>
 filter(is.na(ldb_scores))
```
```{r}
# R functions handle NA automatically
mean(superman$rt_critics_score, na.rm = TRUE)
```

### Converting to Factors for Analysis

For some analyses (like ANOVA), you may want to convert character variables to factors:
```{r}
# Convert for analysis - order levels as needed
superman_for_analysis <- superman |>
 mutate(
    type = factor(type),
    clark_grp = factor(clark_grp),
   height_gap = factor(height_gap, levels = c("Minimal", "Average", "Big"))
 )

# Now ready for ANOVA, contrasts, etc.
levels(superman_for_analysis$height_gap)
```

## Exporting to SPSS

When you export to SPSS, the package automatically:

- Converts categorical variables to numeric codes
- Adds value labels so SPSS displays the category names
- Replaces `NA` with `-99` and marks it as user-defined missing
- Adds variable labels describing each column

### Dataset-Specific Export Functions

Each dataset has a dedicated export function:
```{r eval = FALSE}
# Export individual datasets
export_superman_sav("superman_data.sav")
export_superman_smes_sav("superman_smes_data.sav")
export_hot_ones_sav("hot_ones_data.sav")
export_tip_jokes_sav("tip_jokes_data.sav")
export_mcu_sav("mcu_data.sav")
export_mock_jury_sav("mock_jury_data.sav")
export_candy_sav("candy_data.sav")
export_candy_simple_sav("candy_simple_data.sav")
export_football_sav("football_data.sav")
export_huskers_sav("huskers_data.sav")
export_interpersonal_sav("interpersonal_data.sav")
export_selfdescriptive_sav("selfdescriptive_data.sav")

# Save to a specific location
export_superman_sav("~/Desktop/superman_data.sav")
```

### Generic Export Function

You can also use `export_sav()` with any dataset name:
```{r eval = FALSE}
export_sav("superman", path = "~/Desktop/superman_data.sav")
export_sav("mock_jury") # Saves to working directory
```

### Exporting a Subset of Variables

Select specific columns before exporting:
```{r eval = FALSE}
# Export only selected variables
superman |>
 select(num, media, year, type, clark_height, rt_critics_score) |>
 export_sav(path = "superman_subset.sav")

# Use tidyselect helpers
huskers |>
 select(date, season, opp, result, starts_with("ne_score")) |>
 export_sav(path = "huskers_subset.sav")
```

### Export All Datasets

Create a folder with all `.sav` files at once:
```{r eval = FALSE}
export_all_sav(dir = "~/Desktop/PSYC350_SPSS/")
```

## R vs SPSS: What's Different?

| Aspect | In R | In SPSS (.sav) |
|--------|------|----------------|
| Categorical values | Character strings ("Film", "TV Series") | Numeric codes (1, 2) with value labels |
| Missing values | `NA` | `-99` with missing value definition |
| Variable descriptions | In help files (`?superman`) | Variable labels in Variable View |

## Example Analyses in R

### T-test with Mock Jury Data
```{r}
# Compare sentence length by crime type
# Convert to factor for cleaner output
mock_jury_analysis <- mock_jury |>
 mutate(crime = factor(crime))

t.test(years ~ crime, data = mock_jury_analysis)
```

### ANOVA with Football Data
```{r}
# Compare hippocampus volume across groups
# Character values convert directly to factor
football_analysis <- football |>
 mutate(group = factor(group))

model <- aov(volume ~ group, data = football_analysis)
summary(model)
```

### Correlation with Candy Data
```{r}
# Relationship between sugar content and win percentage
cor.test(candy_simple$sugarpercent, candy_simple$winpercent)
```

### Chi-Square with Tip Jokes Data
```{r}
# Is tipping independent of card type?
# Convert card to factor for labeled output
tip_jokes_analysis <- tip_jokes |>
 mutate(card = factor(card))

table(tip_jokes_analysis$card, tip_jokes_analysis$tip) |>
 chisq.test()
```

## Next Steps

- See `vignette("exporting-spss")` for detailed export options
- See `vignette("dataset-guide")` for complete variable coding schemes
- Use `?dataset_name` to view full documentation for any dataset
- All datasets work with psych350lab analysis functions
