---
title: "Getting Started with psyc350data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with psyc350data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The psyc350data package provides all the datasets you need for PSYC 350 labs. Each dataset is pre-cleaned and ready to use — no downloading files, no import headaches.

## Loading the Package
```{r setup}
library(psyc350data)
```

## Browsing Available Datasets

To see every dataset in the package:
```{r}
list_datasets()
```

For detailed documentation on any dataset, use the help system:
```{r eval = FALSE}
?superman
?hot_ones
?mock_jury
```

## Using Datasets in R

Every dataset is a tibble that's immediately available after loading the package:
```{r}
head(superman)
```
```{r}
str(superman_smes)
```
```{r}
summary(football)
```

### Working with Categorical Variables

Categorical variables are stored as numeric codes in R. The codes map to labels that you can see in the documentation. For example, in the `superman` dataset:

- `type`: 1 = Film, 2 = TV Series, 3 = Serial
- `tomatometer`: 1 = Rotten, 2 = Fresh
- `popular`: 1 = Low, 2 = Mid, 3 = High

You can convert these to factors for analysis or plotting:
```{r}
library(dplyr)

superman |>
  mutate(
    type_label = factor(type, 
                        levels = c(1, 2, 3), 
                        labels = c("Film", "TV Series", "Serial"))
  ) |>
  select(media, year, type, type_label)
```

### Missing Values

In R, missing values are represented as `NA`:
```{r}
# See which Superman entries have missing Letterboxd scores
superman |>
  select(num, media, ldb_likes, ldb_scores) |>
  filter(is.na(ldb_scores))
```

When exported to SPSS, these `NA` values are converted to `-99` and marked as user-defined missing values.

## Example Analyses

### T-test with Mock Jury Data
```{r}
# Compare sentence length by crime type
# crime: 1 = Burglary, 2 = Swindle
t.test(years ~ crime, data = mock_jury)
```

### ANOVA with Football Data
```{r}
# Compare hippocampus volume across groups
# group: 1 = Control, 2 = Football no concussion, 3 = Football with concussion
model <- aov(volume ~ factor(group), data = football)
summary(model)
```

### Correlation with Candy Data
```{r}
# Relationship between sugar content and win percentage
cor.test(candy_simple$sugarpercent, candy_simple$winpercent)
```

### Chi-Square with Tip Jokes Data
```{r}
# Is tipping independent of card type?
table(tip_jokes$card, tip_jokes$tip) |>
  chisq.test()
```

## Next Steps

- See `vignette("exporting-spss")` to learn how to create SPSS files
- Use `?dataset_name` to view full documentation for any dataset
- All datasets work with standard R functions: `t.test()`, `aov()`, `lm()`, `cor.test()`, `chisq.test()`, etc.
