---
title: "Getting Started with psych350data"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Getting Started with psych350data}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>"
)
```

## Overview

The psych350data package provides all the datasets you need for PSYC 350 labs. Each dataset is pre-cleaned and ready to use in two ways:

1. **In R** — Datasets use human-readable values (like "Film", "TV Series") and standard `NA` for missing data. Perfect for analysis with psych350lab functions.

2. **In SPSS** — Export functions create `.sav` files with numeric codes, value labels, and `-99` for missing values.

## Loading the Package
```{r setup}
library(psych350data)
library(dplyr)
```

## Browsing Available Datasets

To see every dataset in the package:
```{r}
list_datasets()
```

For detailed documentation on any dataset, use the help system:
```{r eval = FALSE}
?superman
?hot_ones
?mock_jury
```

## Using Datasets in R

Every dataset is a tibble that's immediately available after loading the package. Categorical variables use descriptive character values that are easy to read and work with:
```{r}
# Categorical variables are human-readable
superman |>
 select(num, media, type, clark_grp, height_gap) |>
 head()
```

### Missing Values in R

Missing values are represented as `NA`, which R functions handle automatically:
```{r}
# See which Superman entries have missing Letterboxd scores
superman |>
 select(num, media, ldb_likes, ldb_scores) |>
 filter(is.na(ldb_scores))
```
```{r}
# R functions handle NA automatically
mean(superman$rt_critics_score, na.rm = TRUE)
```

### Converting to Factors for Analysis

For some analyses (like ANOVA), you may want to convert character variables to factors:
```{r}
# Convert for analysis - order levels as needed
superman_for_analysis <- superman |>
 mutate(
    type = factor(type),
    clark_grp = factor(clark_grp),
   height_gap = factor(height_gap, levels = c("Minimal", "Average", "Big"))
 )

# Now ready for ANOVA, contrasts, etc.
levels(superman_for_analysis$height_gap)
```

## Preparing Data for SPSS-Matched Analysis

If your analysis needs to produce numeric output that matches SPSS — for example, when comparing R results to SPSS output from a lab assignment — use the prep functions to convert categorical variables to the same numeric codes SPSS uses.

### The `prep_data()` Function

`prep_data()` is the main entry point. Pass it a dataset and the dataset name as a string:
```{r}
superman_prepped <- prep_data(superman, "superman")

superman_prepped |>
  select(num, media, type, clark_grp, height_gap) |>
  head()
```

Each character category is replaced with its numeric code. For example, `type` becomes `1` for Film, `2` for TV Series/TV Show, and `3` for Serial — matching the value labels in the SPSS `.sav` file.

### Dataset-Specific Prep Functions

Every dataset also has its own dedicated prep function if you prefer to be explicit:
```{r eval = FALSE}
superman       <- prep_superman(superman)
hot_ones       <- prep_hot_ones(hot_ones)
mcu            <- prep_mcu(mcu)
mock_jury      <- prep_mock_jury(mock_jury)
tip_jokes      <- prep_tip_jokes(tip_jokes)
candy          <- prep_candy(candy)
candy_simple   <- prep_candy_simple(candy_simple)
football       <- prep_football(football)
huskers        <- prep_huskers(huskers)
```

### Keeping Labels for Plotting

By default, prep functions replace character columns with numeric codes and the original strings are lost. If you need the original values for axis labels or legends in plots, use `keep_labels = TRUE`. This retains the original character column alongside the numeric one, with a `_label` suffix:
```{r}
superman_prepped <- prep_superman(superman, keep_labels = TRUE)

# type is now numeric, type_label has the original strings
superman_prepped |>
  select(num, type, type_label) |>
  head()
```

This is useful when you want numeric codes for analysis but readable labels for `ggplot2`:
```{r eval = FALSE}
superman_prepped |>
  ggplot(aes(x = type_label, y = rt_critics_score)) +
  geom_boxplot()
```

### When to Use Prep Functions

| Situation | Recommendation |
|-----------|----------------|
| General R analysis, plotting | Use the raw dataset as-is |
| Matching SPSS numeric output | Use `prep_data()` or a dataset-specific prep function |
| Analysis + plotting with ggplot2 | Use `keep_labels = TRUE` |
| Exporting to `.sav` | Use `export_sav()` — it handles conversion internally |

Note that exporting to SPSS with `export_sav()` does not require running a prep function first — the export functions handle numeric conversion internally.

## Exporting to SPSS

When you export to SPSS, the package automatically:

- Converts categorical variables to numeric codes
- Adds value labels so SPSS displays the category names
- Replaces `NA` with `-99` and marks it as user-defined missing
- Adds variable labels describing each column

### Dataset-Specific Export Functions

Each dataset has a dedicated export function:
```{r eval = FALSE}
# Export individual datasets
export_superman_sav("superman_data.sav")
export_superman_smes_sav("superman_smes_data.sav")
export_hot_ones_sav("hot_ones_data.sav")
export_tip_jokes_sav("tip_jokes_data.sav")
export_mcu_sav("mcu_data.sav")
export_mock_jury_sav("mock_jury_data.sav")
export_candy_sav("candy_data.sav")
export_candy_simple_sav("candy_simple_data.sav")
export_football_sav("football_data.sav")
export_huskers_sav("huskers_data.sav")
export_interpersonal_sav("interpersonal_data.sav")
export_selfdescriptive_sav("selfdescriptive_data.sav")

# Save to a specific location
export_superman_sav("data/superman_data.sav")
```

### Generic Export Function

You can also use `export_sav()` with any dataset name:
```{r eval = FALSE}
export_sav("superman", path = "data/superman_data.sav")
export_sav("mock_jury") # Saves to working directory
```

### Exporting a Subset of Variables

Select specific columns before exporting:
```{r eval = FALSE}
# Export only selected variables
superman |>
 select(num, media, year, type, clark_height, rt_critics_score) |>
 export_sav(path = "superman_subset.sav")

# Use tidyselect helpers
huskers |>
 select(date, season, opp, result, starts_with("ne_score")) |>
 export_sav(path = "huskers_subset.sav")
```

### Export All Datasets

Create a folder with all `.sav` files at once:
```{r eval = FALSE}
export_all_sav(dir = "data/PSYC350_SPSS/")
```

## R vs SPSS: What's Different?

| Aspect | In R | In SPSS (.sav) |
|--------|------|----------------|
| Categorical values | Character strings ("Film", "TV Series") | Numeric codes (1, 2) with value labels |
| Missing values | `NA` | `-99` with missing value definition |
| Variable descriptions | In help files (`?superman`) | Variable labels in Variable View |
