---
title: "Dataset Reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dataset Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(psych350data)
library(dplyr)
```

A complete reference for all datasets in the psych350data package. Each entry includes a description, source attribution, variable overview, and the categorical variable codes used in SPSS exports and after `prep_*()`.

For guidance on how to use these datasets, see the [Getting Started](getting-started.qmd) vignette.

## Quick Reference

+-------------------------+---------------------------------+-------------------------------------+
| Dataset                 | Related Functions               | Source                              |
+=========================+=================================+=====================================+
| `superman`              | `prep_superman()`               | Rotten Tomatoes, Letterboxd, IMDb   |
|                         |                                 |                                     |
|                         | `label_superman()`              |                                     |
|                         |                                 |                                     |
|                         | `export_superman_sav()`         |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `superman_smes`         | `prep_superman_smes()`          | Simulated                           |
|                         |                                 |                                     |
|                         | `label_superman_smes()`         |                                     |
|                         |                                 |                                     |
|                         | `export_superman_smes_sav()`    |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `superman_movies`       | `prep_superman_movies()`        | IMDb Box Office Mojo                |
|                         |                                 |                                     |
|                         | `label_superman_movies`         |                                     |
|                         |                                 |                                     |
|                         | `export_superman_movies_sav()`  |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `hotones`               | `prep_hotones()`                | Hot Ones / First We Feast (YouTube) |
|                         |                                 |                                     |
|                         | `label_hotones()`               |                                     |
|                         |                                 |                                     |
|                         | `export_hotones_sav()`          |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `hotones_sauces`        | `prep_hotones_sauces()`         | Hot Ones / First We Feast (YouTube) |
|                         |                                 |                                     |
|                         | `label_hotones_sauces()`        |                                     |
|                         |                                 |                                     |
|                         | `export_hotones_sauces_sav()`   |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `hotones_episodes`      | `prep_hotones_episodes()`       | Hot Ones / First We Feast (YouTube) |
|                         |                                 |                                     |
|                         | `label_hotones_episodes()`      |                                     |
|                         |                                 |                                     |
|                         | `export_hotones_episodes_sav()` |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `tip_jokes`             | `prep_tip_jokes()`              | Gueguen (2002)                      |
|                         |                                 |                                     |
|                         | `label_tip_jokes()`             |                                     |
|                         |                                 |                                     |
|                         | `export_tip_jokes_sav()`        |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `mcu`                   | `prep_mcu()`                    | IMDb via openintro package          |
|                         |                                 |                                     |
|                         | `label_mcu()`                   |                                     |
|                         |                                 |                                     |
|                         | `export_mcu_sav()`              |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `mock_jury`             | `prep_mock_jury()`              | Plaster (1989)                      |
|                         |                                 |                                     |
|                         | `label_mock_jury()`             |                                     |
|                         |                                 |                                     |
|                         | `export_mock_jury_sav()`        |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `candy`                 | `prep_candy()`                  | FiveThirtyEight                     |
|                         |                                 |                                     |
|                         | `label_candy()`                 |                                     |
|                         |                                 |                                     |
|                         | `export_candy_sav()`            |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `candy_simple`          | `prep_candy_simple()`           | FiveThirtyEight                     |
|                         |                                 |                                     |
|                         | `label_candy_simple()`          |                                     |
|                         |                                 |                                     |
|                         | `export_candy_simple_sav()`     |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `football`              | `prep_football()`               | Singh et al. (2014), JAMA           |
|                         |                                 |                                     |
|                         | `label_football()`              |                                     |
|                         |                                 |                                     |
|                         | `export_football_sav()`         |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `huskers`               | `prep_huskers()`                | Historical records                  |
|                         |                                 |                                     |
|                         | `label_huskers()`               |                                     |
|                         |                                 |                                     |
|                         | `export_huskers_sav()`          |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `cheese_data`           | `prep_cheese()`                 | cheese.com via TidyTuesday          |
|                         |                                 |                                     |
|                         | `label_cheese()`                |                                     |
|                         |                                 |                                     |
|                         | `export_cheese_sav()`           |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `lpd_data`              | `prep_lpd()`                    | Lincoln Police Dept. Open Data      |
|                         |                                 |                                     |
|                         | `label_lpd()`                   |                                     |
|                         |                                 |                                     |
|                         | `export_lpd_sav()`              |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `parent_child_data`     | `prep_parent_child()`           | Simulated                           |
|                         |                                 |                                     |
|                         | `label_parent_child()`          |                                     |
|                         |                                 |                                     |
|                         | `export_parent_child_sav()`     |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `hindsight_mg_data`     | `prep_hindsight_bg()`[^1]       | Simulated                           |
|                         |                                 |                                     |
|                         | `label_hindsight_mg()`          |                                     |
|                         |                                 |                                     |
|                         | `export_hindsight_mg_sav()`     |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `hindsight_wg_data`     | `prep_hindsight_wg()`           | Simulated                           |
|                         |                                 |                                     |
|                         | `label_hindsight_wg()`          |                                     |
|                         |                                 |                                     |
|                         | `export_hindsight_wg_sav()`     |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `interpersonal_data`    | `prep_interpersonal()`          | Simulated                           |
|                         |                                 |                                     |
|                         | `label_interpersonal()`         |                                     |
|                         |                                 |                                     |
|                         | `export_interpersonal_sav()`    |                                     |
+-------------------------+---------------------------------+-------------------------------------+
| `self_descriptive_data` | `prep_self_descriptive()`       | Simulated                           |
|                         |                                 |                                     |
|                         | `label_self_descriptive()`      |                                     |
|                         |                                 |                                     |
|                         | `export_selfdescriptive_sav()`  |                                     |
+-------------------------+---------------------------------+-------------------------------------+

[^1]: The prep function uses `bg` (between-groups) while the dataset and export use `mg`.

::: callout-note
A **combined** Superman dataset (`superman_combined`) can be exported with `export_superman_combined_sav()`. It joins `superman_movies` with `superman` actor data using `join_superman_data()`.
:::

Use `list_datasets()` to see all available dataset names from the R console.

## How Categorical Variables Work

**In R (raw):** Categorical variables use human-readable character values like `"Film"`, `"Control"`, or `"Minimal"`. These are ideal for plotting and exploration.

**After `prep_*()`:** Character categories are replaced with numeric codes that match SPSS. Use this when your R output needs to match SPSS output or when working with psych350lab functions.

**In SPSS exports:** Numeric codes are stored with value labels, so SPSS displays both the number and the category name.

```{r}
# Raw R data: human-readable categories
superman |>
  select(num, media, type, age_grp) |>
  head(4)
```

```{r}
# After prep: numeric codes matching SPSS
prep_superman(superman) |>
  select(num, media, type, age_grp) |>
  head(4)
```

------------------------------------------------------------------------

## Superman Actor Data {#sec-superman}

**Dataset:** `superman` \| **11 rows, 27 columns**

Physical characteristics and ratings data for actors who have played Superman across film, TV, and serial media. Includes actor heights, Rotten Tomatoes scores, Letterboxd ratings, and popularity metrics.

**Source:** Compiled from Rotten Tomatoes, Letterboxd, and IMDb.

```{r}
glimpse(superman)
```

### Categorical Variables (SPSS codes)

| Variable      | Values                              |
|---------------|-------------------------------------|
| `type`        | 1 = Film, 2 = TV Series, 3 = Serial |
| `clark_grp`   | 1 = Under 6ft, 2 = 6ft or taller    |
| `height_gap`  | 1 = Minimal, 2 = Average, 3 = Big   |
| `age_grp`     | 1 = Minimal, 2 = Average, 3 = Big   |
| `tomatometer` | 1 = Rotten, 2 = Fresh               |
| `popular`     | 1 = Low, 2 = Mid, 3 = High          |

------------------------------------------------------------------------

## Superman SMES Data {#sec-superman-smes}

**Dataset:** `superman_smes` \| **47 rows, 7 columns**

Simulated participant ratings on the Subjective Media Experience Scale (SMES), grouped by both the height gap and age difference between the Superman and Lois Lane actors. The `emotion` variable requires data prep conversion to factor; all other variables are already numeric.

**Source:** Simulated data for teaching purposes.

```{r}
glimpse(superman_smes)
```

### Categorical Variables (SPSS codes)

| Variable     | Values                            |
|--------------|-----------------------------------|
| `height_gap` | 1 = Minimal, 2 = Average, 3 = Big |
| `age_grp`    | 1 = Minimal, 2 = Average, 3 = Big |
| `emotion`    | 1 = Fear, 2 = Joy, 3 = Sadness, 4 = Anger, 5 = Disgust, 6 = Anxiety |

### Scale Information

| Subscale               | Range |
|------------------------|-------|
| `emotional_impact`     | 4–20  |
| `aesthetic_appeal`     | 3–15  |
| `cognitive_engagement` | 0–7   |

------------------------------------------------------------------------

## Superman Movies Data {#sec-superman-movies}

**Dataset:** `superman_movies` \| **10 rows, 22 columns**

Box office and production data for Superman theatrical films, including budget, domestic and international grosses, and MPAA ratings.

**Source:** IMDb Box Office Mojo.

```{r}
glimpse(superman_movies)
```

### Categorical Variables (SPSS codes)

| Variable         | Values                                        |
|------------------|-----------------------------------------------|
| `mpaa`           | 1 = G, 2 = PG, 3 = PG-13, 4 = R               |
| `budget_cat`     | 1 = Low, 2 = Medium, 3 = High (tercile-based) |
| `box_office_cat` | 1 = Low, 2 = Medium, 3 = High (tercile-based) |

------------------------------------------------------------------------

## Hot Ones Guest Data {#sec-hotones}

**Dataset:** `hotones` \| **Export:** `export_hotones_sav()`

Data on guests from the YouTube show *Hot Ones*, including demographic information, Scoville ratings for each sauce consumed (SHU_1 through SHU_10), and whether the guest succeeded in eating all ten wings.

**Source:** [Hot Ones / First We Feast](https://www.youtube.com/firstwefeast) (YouTube).

```{r}
glimpse(hotones)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----|----|
| `gender` | 1 = Male, 2 = Female |
| `result` | 1 = Succeeded, 2 = Failed, 3 = Incomplete |
| `occupation` | 1 = Rapper, 2 = Athlete, 3 = Actor, 4 = Actor-Comedian, 5 = Comedian, 6 = Chef, 7 = Actor-Musician, 8 = Musician, 9 = DJ, 10 = YouTuber, 11 = Model, 12 = Wrestler, 13 = Magician, 14 = Other |

------------------------------------------------------------------------

## Hot Ones Sauces Data {#sec-hotones-sauces}

**Dataset:** `hotones_sauces` \| **Export:** `export_hotones_sauces_sav()`

Data on the hot sauces used in each season and position of Hot Ones, including Scoville Heat Unit (SHU) ratings. All variables are numeric — no categorical conversion needed.

**Source:** [Hot Ones / First We Feast](https://www.youtube.com/firstwefeast) (YouTube).

```{r}
glimpse(hotones_sauces)
```

------------------------------------------------------------------------

## Hot Ones Episodes Data {#sec-hotones-episodes}

**Dataset:** `hotones_episodes` \| **Export:** `export_hotones_episodes_sav()`

Episode-level data from Hot Ones including guest names, episode titles, and YouTube engagement metrics (views, likes, comments). All variables are numeric or character — no categorical conversion needed.

**Source:** [Hot Ones / First We Feast](https://www.youtube.com/firstwefeast) (YouTube).

```{r}
glimpse(hotones_episodes)
```

------------------------------------------------------------------------

## Tip-Jokes Experiment Data {#sec-tip-jokes}

**Dataset:** `tip_jokes` \| **211 rows, 5 columns**

Experimental data examining whether a waiter leaving a joke or an advertisement on a card affects tipping behavior. All variables are already stored as numeric codes.

**Source:** Gueguen, N. (2002). The effects of a joke on tipping when it is delivered at the same time as the bill. *Journal of Applied Social Psychology*, 32(9), 1955–1963.

```{r}
glimpse(tip_jokes)
```

### Variables (SPSS codes)

| Variable | Values                                   |
|----------|------------------------------------------|
| `card`   | 1 = Advertisement, 2 = Joke, 3 = None    |
| `tip`    | 0 = No, 1 = Yes                          |
| `ad`     | 0 = No, 1 = Yes (received advertisement) |
| `joke`   | 0 = No, 1 = Yes (received joke)          |
| `none`   | 0 = No, 1 = Yes (no card)                |

------------------------------------------------------------------------

## MCU Films Data {#sec-mcu}

**Dataset:** `mcu` \| **23 rows, 11 columns**

Box office performance and Rotten Tomatoes scores for Marvel Cinematic Universe films through the Infinity Saga (Phases 1–3).

**Source:** Internet Movie Database (IMDb), via the [openintro](https://github.com/OpenIntroStat/openintro-statistics) package.

```{r}
glimpse(mcu)
```

### Categorical Variables (SPSS codes)

| Variable | Values                                |
|----------|---------------------------------------|
| `phase`  | 1 = Phase 1, 2 = Phase 2, 3 = Phase 3 |
| `favor`  | 1 = Critics, 2 = Audience             |

------------------------------------------------------------------------

## Mock Jury Data {#sec-mock-jury}

**Dataset:** `mock_jury` \| **114 rows, 17 columns**

Data from a study examining the effects of defendant physical attractiveness on mock jury sentencing decisions. Participants rated defendants of varying attractiveness levels who were charged with either burglary or swindle. All categorical variables are already stored as numeric codes.

**Source:** Plaster, M. E. (1989). *The effects of physical attractiveness on mock jury decisions*. Unpublished manuscript, East Carolina University.

```{r}
glimpse(mock_jury)
```

### Categorical Variables (SPSS codes)

| Variable | Values                                       |
|----------|----------------------------------------------|
| `attr`   | 1 = Beautiful, 2 = Average, 3 = Unattractive |
| `crime`  | 1 = Burglary, 2 = Swindle                    |

------------------------------------------------------------------------

## Candy Rankings Data {#sec-candy}

**Full dataset:** `candy` \| **85 rows, 13 columns** **Simplified:** `candy_simple` \| **85 rows, 5 columns**

Candy power rankings based on 269,000 head-to-head matchups. The full dataset includes nine binary ingredient indicators, plus sugar percentile, price percentile, and win percentage. The simplified version includes only `chocolate`, `sugarpercent`, `pricepercent`, and `winpercent`.

**Source:** [The Ultimate Halloween Candy Power Ranking](https://web.archive.org/web/20201101001713/https://fivethirtyeight.com/videos/the-ultimate-halloween-candy-power-ranking/), FiveThirtyEight (2017).

```{r}
glimpse(candy_simple)
```

### Binary Variables (SPSS codes)

All binary ingredient variables: 0 = No, 1 = Yes.

Variables in the full dataset: `chocolate`, `fruity`, `caramel`, `peanutyalmondy`, `nougat`, `crispedricewafer`, `hard`, `bar`, `pluribus`.

The simplified dataset contains only `chocolate`.

------------------------------------------------------------------------

## Football Concussion Data {#sec-football}

**Dataset:** `football` \| **75 rows, 3 columns**

Brain volume measurements comparing football players (with and without concussion history) to non-football-playing controls.

**Source:** Singh, R., Meier, T. B., Kuplicki, R., Savitz, J., Mukai, I., Cavanagh, L., Allen, T., Teague, T. K., Nerio, C., Polanski, D., & Bellgowan, P. S. F. (2014). Relationship of collegiate football experience and concussion with hippocampal volume and cognitive outcomes. *JAMA*, 311(18), 1883–1888.

```{r}
glimpse(football)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----|----|
| `group` | 1 = Control, 2 = Football no concussion, 3 = Football with concussion |

------------------------------------------------------------------------

## Nebraska Football Box Scores {#sec-huskers}

**Dataset:** `huskers` \| **Export:** `export_huskers_sav()`

Game-level team statistics for all Nebraska Cornhuskers football games from September 1962 through the 2024 season. Includes scoring, rushing, passing, turnovers, penalties, point spreads, and weather data. This is a large dataset with 68 columns.

**Source:** Compiled from historical Nebraska football records. Weather data sourced from DarkSky API and Weather Underground.

```{r}
glimpse(huskers)
```

### Categorical Variables (SPSS codes)

| Variable     | Values                                                     |
|--------------|------------------------------------------------------------|
| `result`     | 1 = Win, 2 = Loss, 3 = Tie                                 |
| `site`       | 1 = Home, 2 = Away, 3 = Neutral (home), 4 = Neutral (away) |
| `conference` | 0 = Non-conference, 1 = Conference                         |

------------------------------------------------------------------------

## Cheese Characteristics Data {#sec-cheese}

**Dataset:** `cheese_data` \| **Export:** `export_cheese_sav()`

Cleaned subset of the cheese.com dataset originally featured in TidyTuesday (June 2024). Contains cheeses with known calcium content, with fat content and milk source cleaned and recoded for teaching data entry and transformation.

**Source:** [cheese.com](https://www.cheese.com/) via [TidyTuesday 2024-06-04](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-06-04/readme.md).

```{r}
glimpse(cheese_data)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----|----|
| `milk_source` | 1 = Cow, 2 = Goat, 3 = Sheep, 4 = Buffalo, 5 = Multiple, 6 = Other |
| `vegetarian` | 0 = No, 1 = Yes |

------------------------------------------------------------------------

## Lincoln Police Department Traffic Stops {#sec-lpd}

**Dataset:** `lpd_data` \| **Export:** `export_lpd_sav()`

Traffic stop records from the Lincoln Police Department, compiled from a multi-sheet Excel file with one sheet per year and stacked into a single data frame. Date and time variables have been parsed and a time-of-day category added. All categorical variables are stored as integers with SPSS-style value labels.

**Source:** [Lincoln Police Department Open Data Portal](https://opendata.lincoln.ne.gov/datasets/lpd-traffic-stops-2023/about).

```{r}
glimpse(lpd_data)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----|----|
| `time_of_day` | 1 = Morning (5am–noon), 2 = Afternoon (noon–5pm), 3 = Evening (5pm–9pm), 4 = Night (9pm–5am) |
| `race` | 1 = White, 2 = Black or African American, 3 = Hispanic or Latino, 4 = Asian, 5 = American Indian or Alaska Native, 6 = Native Hawaiian or Pacific Islander, 7 = Two or More Races, 8 = Unknown |
| `sex` | 1 = Male, 2 = Female, 3 = Unknown |
| `reason` | 1 = Traffic probable cause, 2 = Criminal probable cause, 3 = Other |
| `outcome` | 1 = Traffic warning, 2 = Traffic official, 3 = Criminal cite & release, 4 = Lodged in jail, 5 = None |
| `search` | 1 = None, 2 = Incident to arrest, 3 = Inventory, 4 = Consent, 5 = Probable cause |

------------------------------------------------------------------------

## Parent-Child Observation Data {#sec-parent-child}

**Dataset:** `parent_child_data` \| **100 rows, 11 columns**

Simulated observational data from a study examining the effects of a remediation treatment on parenting behaviours. Includes maternal and child demographics alongside coded behavioural counts (praise, directives, and negative behaviours) from structured observations. All categorical variables are already stored as numeric codes.

**Source:** Simulated data generated to resemble a plausible parent-child observational study.

```{r}
glimpse(parent_child_data)
```

### Categorical Variables (SPSS codes)

| Variable    | Values                                       |
|-------------|----------------------------------------------|
| `magegroup` | 1 = 18–27, 2 = 28–35                         |
| `cagegroup` | 1 = 2–3 years, 2 = 4–5 years                 |
| `famtype`   | 1 = 2-parent family, 2 = Mother-only         |
| `CLINREM`   | 0 = Remediation suggested, 1 = Not suggested |
| `tx`        | 0 = Control, 1 = Remediation                 |

------------------------------------------------------------------------

## Hindsight Bias Data — Between-Groups {#sec-hindsight-mg}

**Dataset:** `hindsight_mg_data` \| **600 rows, 7 columns**

Simulated data from a hindsight bias study in which participants viewed celebrity faces and estimated recognition time at baseline and again after receiving an outcome cue. The between-groups factor is whether participants saw old or new faces in the hindsight phase. Each row represents one participant-face trial (60 participants × 10 faces). All categorical variables are already stored as numeric codes.

**Source:** Simulated data generated to illustrate hindsight bias in a between-groups design.

```{r}
glimpse(hindsight_mg_data)
```

### Categorical Variables (SPSS codes)

| Variable     | Values                                      |
|--------------|---------------------------------------------|
| `condition`  | 1 = Old, 2 = New                            |
| `fame_level` | 1 = Extremely Famous, 2 = Moderately Famous |
| `correct`    | 0 = Incorrect, 1 = Correct                  |

------------------------------------------------------------------------

## Hindsight Bias Data — Within-Groups {#sec-hindsight-wg}

**Dataset:** `hindsight_wg_data` \| **60 rows, 6 columns**

Participant-level summary of the hindsight bias study, averaged across faces within each fame level. Each row is one participant with separate columns for extremely and moderately famous faces at baseline and hindsight phases. Suitable for a 2×2 repeated-measures or mixed ANOVA. All categorical variables are already stored as numeric codes.

**Source:** Simulated data generated to illustrate hindsight bias in a within-groups design.

```{r}
glimpse(hindsight_wg_data)
```

### Categorical Variables (SPSS codes)

| Variable    | Values           |
|-------------|------------------|
| `condition` | 1 = Old, 2 = New |

------------------------------------------------------------------------

## Interpersonal Relationships Survey Data {#sec-interpersonal}

**Dataset:** `interpersonal_data` \| **574 rows, 33 columns**

Simulated survey data from college students at a predominantly white rural state university. Contains demographics, relationship variables, and scores on several interpersonal and psychological scales. All categorical variables are already stored as numeric codes.

**Source:** Simulated data generated to resemble plausible survey responses from undergraduate psychology students.

```{r}
glimpse(interpersonal_data)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----|----|
| `gender` | 1 = Male, 2 = Female, 3 = Another |
| `race` | 1 = Asian, 2 = Black, 3 = Indigenous, 4 = Latino/Hispanic, 5 = Middle Eastern, 6 = White, 7 = Other |
| `hand` | 1 = Right, 2 = Left, 3 = Both |
| `community` | 1 = Rural, 2 = Small town, 3 = Suburban, 4 = Urban |
| `famclass` | 1 = Working, 2 = Lower, 3 = Lower middle, 4 = Upper middle, 5 = Upper |
| `greek_in` | 1 = Independent, 2 = Greek |

### Scale Variables

The following composite scale scores are included: DAQ (Dating and Assertion Questionnaire) subscales, IDI (Interpersonal Dependency Inventory) subscales, IRI (Interpersonal Reactivity Index) subscales, Sociability subscales, RISC (Relational-Interdependent Self-Construal), GCB (Generic Conspiracist Beliefs), and LSAS-SR (Liebowitz Social Anxiety Scale — Self Report).

------------------------------------------------------------------------

## Self-Descriptive Survey Data {#sec-self-descriptive}

**Dataset:** `self_descriptive_data` \| **547 rows, 37 columns**

Simulated survey data from college students at a predominantly white rural state university. Contains demographics, relationship variables, and scores on personality and self-concept measures. All categorical variables are already stored as numeric codes.

**Source:** Simulated data generated to resemble plausible survey responses from undergraduate psychology students.

```{r}
glimpse(self_descriptive_data)
```

### Categorical Variables

Same demographic coding as `interpersonal_data` (see above).

### Scale Variables

The following composite scale scores are included: TIPI (Ten-Item Personality Inventory) Big Five subscales, MAAS (Mindful Attention Awareness Scale) subscales, RFQ (Regulatory Focus Questionnaire) subscales, ATQ-30 (Automatic Thoughts Questionnaire) subscales, RSE (Rosenberg Self-Esteem), and NGSE (New General Self-Efficacy).
