---
title: "Dataset Reference Guide"
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(psych350data)
library(dplyr)
```

A quick overview of datasets in the psych350data package.

## Quick Reference

| Dataset | Labeling & Prep Functions | Export Function |
|---------------------------|---------------------------|-------------------|
| `superman` | `label_superman()` | `export_superman_sav()` |
| `superman_smes` | `label_superman_smes()` | `export_superman_smes_sav()` |
| `superman_movies` | `label_superman_movies()` | `export_superman_movies_sav()` |
| `hotones` | `label_hotones()` | `export_hotones_sav()` |
| `hotones_sauces` | `label_hotones_sauces()` | `export_hotones_sauces_sav()` |
| `hotones_episodes` | `label_hotones_episodes` | `export_hotones_episodes_sav()` |
| `tip_jokes` | `tip_jokes` | `export_tip_jokes_sav()` |
| `mcu` | `label_mcu` | `export_mcu_sav()` |
| `mock_jury` | `label_mock_jury()` | `export_mock_jury_sav()` |
| `candy` | `label_candy()` | `export_candy_sav()` |
| `candy_simple` | `label_candy_simpl()` | `export_candy_simple_sav()` |
| `football` | `football()` | `export_football_sav()` |
| `huskers` | `label_husker()` | `export_huskers_sav()` |
| `lpd_data` | `label_lpd()` | `export_lpd_sav()` |
| `parent_child` | `label_parent_chil()` | `export_parent_child_sav()` |
| `cheese` | `label_cheese()` | `export_cheese_sav()` |
| `interpersonal_data` | `label_interpersonal_data()` | `export_interpersonal_sav()` |
| `self_descriptive_data` | `labelself_descriptive_data()` | `export_selfdescriptive_sav()` |

Use `list_datasets()` to see all available datasets.

## How Categorical Variables Work

**In R:** Categorical variables use human-readable character values (e.g., `"Film"`, `"TV Series"`, `"Minimal"`).

**In SPSS exports:** These are automatically converted to numeric codes with value labels.

```{r}
# Example: categorical variables are human-readable in R
superman |>
  select(num, media, type, height_gap) |>
  head(4)
```

------------------------------------------------------------------------

## Superman Actor Data

**Dataset:** `superman` \| **Export:** `export_superman_sav()`

Data on actors who have portrayed Clark Kent/Superman across film, TV, and serial media.

```{r}
glimpse(superman)
```

### Categorical Variables (SPSS codes)

| Variable      | Values                              |
|---------------|-------------------------------------|
| `type`        | 1 = Film, 2 = TV Series, 3 = Serial |
| `clark_grp`   | 1 = Under 6ft, 2 = 6ft or taller    |
| `height_gap`  | 1 = Minimal, 2 = Average, 3 = Big   |
| `tomatometer` | 1 = Rotten, 2 = Fresh               |
| `popular`     | 1 = Low, 2 = Mid, 3 = High          |

------------------------------------------------------------------------

## Superman SMES Data

**Dataset:** `superman_smes` \| **Export:** `export_superman_smes_sav()`

Simulated participant ratings on the Subjective Media Experience Scale (SMES).

```{r}
glimpse(superman_smes)
```

### Scale Information

| Subscale               | Range |
|------------------------|-------|
| `emotional_impact`     | 4-20  |
| `aesthetic_appeal`     | 3-15  |
| `cognitive_engagement` | 0-7   |

------------------------------------------------------------------------

## Superman Movies Data

**Dataset:** `superman_movies` \| **Export:** `export_superman_movies_sav()`

Box office and production data for Superman theatrical films.

```{r}
glimpse(superman_movies)
```

### Categorical Variables (SPSS codes)

| Variable         | Values                          |
|------------------|---------------------------------|
| `mpaa`           | 1 = G, 2 = PG, 3 = PG-13, 4 = R |
| `budget_cat`     | 1 = Low, 2 = Medium, 3 = High   |
| `box_office_cat` | 1 = Low, 2 = Medium, 3 = High   |

------------------------------------------------------------------------

## Hot Ones Guest Data

**Dataset:** `hotones` \| **Export:** `export_hotones_sav()`

Data on guests from the YouTube show Hot Ones.

```{r}
glimpse(hotones)
```

### Categorical Variables (SPSS codes)

| Variable     | Values                                    |
|--------------|-------------------------------------------|
| `gender`     | 1 = Male, 2 = Female                      |
| `result`     | 1 = Succeeded, 2 = Failed, 3 = Incomplete |
| `occupation` | 1-14 (Rapper, Athlete, Actor, etc.)       |

------------------------------------------------------------------------

## Hot Ones Sauces Data

**Dataset:** `hotones_sauces` \| **Export:** `export_hotones_sauces_sav()`

Hot sauce data by season and position in the lineup.

```{r}
glimpse(hotones_sauces)
```

------------------------------------------------------------------------

## Hot Ones Episodes Data

**Dataset:** `hotones_episodes` \| **Export:** `export_hotones_episodes_sav()`

Episode-level data with YouTube engagement metrics.

```{r}
glimpse(hotones_episodes)
```

------------------------------------------------------------------------

## Tip-Jokes Experiment Data

**Dataset:** `tip_jokes` \| **Export:** `export_tip_jokes_sav()`

Experimental data from Gueguen (2002) examining whether a waiter leaving a joke or advertisement on a card affects tipping.

```{r}
glimpse(tip_jokes)
```

### Categorical Variables (SPSS codes)

| Variable | Values                                |
|----------|---------------------------------------|
| `card`   | 1 = Advertisement, 2 = Joke, 3 = None |
| `tip`    | 0 = No, 1 = Yes                       |

------------------------------------------------------------------------

## MCU Films Data

**Dataset:** `mcu` \| **Export:** `export_mcu_sav()`

Marvel Cinematic Universe films through the Infinity Saga. Credit to (openstats)\[https://github.com/OpenIntroStat/openintro-statistics\]

```{r}
glimpse(mcu)
```

### Categorical Variables (SPSS codes)

| Variable | Values                                |
|----------|---------------------------------------|
| `phase`  | 1 = Phase 1, 2 = Phase 2, 3 = Phase 3 |
| `favor`  | 1 = Critics, 2 = Audience             |

------------------------------------------------------------------------

## Mock Jury Data

**Dataset:** `mock_jury` \| **Export:** `export_mock_jury_sav()`

Data from Plaster (1989) on effects of defendant attractiveness on mock jury sentencing.

```{r}
glimpse(mock_jury)
```

### Categorical Variables (SPSS codes)

| Variable | Values                                       |
|----------|----------------------------------------------|
| `attr`   | 1 = Beautiful, 2 = Average, 3 = Unattractive |
| `crime`  | 1 = Burglary, 2 = Swindle                    |

------------------------------------------------------------------------

## Candy Rankings Data

**Full dataset:** `candy` \| **Export:** `export_candy_sav()` **Simplified:** `candy_simple` \| **Export:** `export_candy_simple_sav()`

Candy power rankings based on 269,000 head-to-head matchups from [The Ultimate Halloween Candy Power Ranking (2017)](https://web.archive.org/web/20201101001713/https://fivethirtyeight.com/videos/the-ultimate-halloween-candy-power-ranking/)

```{r}
glimpse(candy_simple)
```

### Binary Variables (SPSS codes)

All binary variables: 0 = No, 1 = Yes

Variables: `chocolate`, `fruity`, `caramel`, `peanutyalmondy`, `nougat`, `crispedricewafer`, `hard`, `bar`, `pluribus`

------------------------------------------------------------------------

## Football Concussion Data

**Dataset:** `football` \| **Export:** `export_football_sav()`

Brain measurements comparing football players to controls (Singh et al., 2014).

```{r}
glimpse(football)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------------------------------------|--------------------------------|
| `group` | 1 = Control, 2 = Football no concussion, 3 = Football with concussion |

------------------------------------------------------------------------

## Nebraska Football Box Scores

**Dataset:** `huskers` \| **Export:** `export_huskers_sav()`

Game-level statistics for Nebraska Cornhuskers football (1962-2024).

```{r}
glimpse(huskers)
```

### Categorical Variables (SPSS codes)

| Variable     | Values                                                     |
|--------------|------------------------------------------------------------|
| `result`     | 1 = Win, 2 = Loss, 3 = Tie                                 |
| `site`       | 1 = Home, 2 = Away, 3 = Neutral (home), 4 = Neutral (away) |
| `conference` | 0 = Non-conference, 1 = Conference                         |

------------------------------------------------------------------------

## Interpersonal Relationships Survey Data

**Dataset:** `interpersonal_data` \| **Export:** `export_interpersonal_sav()`

Simulated survey data (n = 574) with demographics and interpersonal scales.

```{r}
glimpse(interpersonal_data)
```

### Categorical Variables (SPSS codes)

| Variable | Values |
|----------------------------------------|--------------------------------|
| `gender` | 1 = Male, 2 = Female, 3 = Another |
| `race` | 1 = Asian, 2 = Black, 3 = Indigenous, 4 = Latino/Hispanic, 5 = Middle Eastern, 6 = White, 7 = Other |
| `hand` | 1 = Right, 2 = Left, 3 = Both |
| `community` | 1 = Rural, 2 = Small town, 3 = Suburban, 4 = Urban |
| `famclass` | 1 = Working, 2 = Lower, 3 = Lower middle, 4 = Upper middle, 5 = Upper |
| `greek_in` | 1 = Independent, 2 = Greek |

### Scale Variables

DAQ subscales, IDI subscales, IRI subscales, Sociability subscales, RISC, GCB, LSAS-SR

------------------------------------------------------------------------

## Self-Descriptive Survey Data

**Dataset:** `self_descriptive_data` \| **Export:** `export_selfdescriptive_sav()`

Simulated survey data (n = 547) with demographics and personality scales.

```{r}
glimpse(self_descriptive_data)
```

### Categorical Variables

Same coding as `interpersonal_data` for demographic variables.

### Scale Variables

TIPI Big Five, MAAS subscales, RFQ subscales, ATQ-30 subscales, Rosenberg Self-Esteem, NGSE

------------------------------------------------------------------------
