---
title: "Exporting Datasets to SPSS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exporting Datasets to SPSS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

All psych350data datasets can be exported as SPSS `.sav` files with full metadata. You do **not** need to prep or transform the data first — the export functions handle everything internally. Each export automatically:

- Converts categorical variables from character strings to numeric codes
- Adds **value labels** so SPSS displays category names in Data View
- Adds **variable labels** describing each column in Variable View
- Replaces `NA` with **`-99`** and registers it as user-defined missing


## Setup
```{r setup}
library(psych350data)
library(dplyr)
```


## Quick Export

Each dataset has a dedicated export function that writes a `.sav` file:
```{r}
#| eval: false
# Export to current working directory (uses default filename)
export_superman_sav()

# Export to a specific location
export_football_sav("~/Desktop/football_data.sav")
```

Or use the generic `export_sav()` function with a dataset name:
```{r}
#| eval: false
export_sav("superman", path = "superman_data.sav")
export_sav("hotones")
```


## All Export Functions

| Function | Default filename |
|----------|-----------------|
| `export_superman_sav()` | `superman_data.sav` |
| `export_superman_smes_sav()` | `superman_smes_data.sav` |
| `export_superman_movies_sav()` | `superman_movies_data.sav` |
| `export_superman_combined_sav()` | `superman_combined_data.sav` |
| `export_hotones_sav()` | `hotones_data.sav` |
| `export_hotones_sauces_sav()` | `hotones_sauces_data.sav` |
| `export_hotones_episodes_sav()` | `hotones_episodes_data.sav` |
| `export_tip_jokes_sav()` | `tip_jokes_data.sav` |
| `export_mcu_sav()` | `mcu_data.sav` |
| `export_mock_jury_sav()` | `mock_jury_data.sav` |
| `export_candy_sav()` | `candy_data.sav` |
| `export_candy_simple_sav()` | `candy_simple_data.sav` |
| `export_football_sav()` | `football_data.sav` |
| `export_huskers_sav()` | `huskers_data.sav` |
| `export_cheese_sav()` | `cheese_data.sav` |
| `export_lpd_sav()` | `lpd_data.sav` |
| `export_parent_child_sav()` | `parent_child_data.sav` |
| `export_hindsight_mg_sav()` | `hindsight_mg_data.sav` |
| `export_hindsight_wg_sav()` | `hindsight_wg_data.sav` |
| `export_interpersonal_sav()` | `interpersonal_data.sav` |
| `export_selfdescriptive_sav()` | `selfdescriptive_data.sav` |


## Exporting for Answer Keys vs. Student Data {#sec-answer-key-vs-student}

A common instructor workflow is to export the **full dataset** as an answer key and a **subset of variables** as the version students receive.

### Full Dataset (Answer Key)

The default export includes every variable with all labels:
```{r}
#| eval: false
# Full dataset — all variables, all labels, -99 for missing
export_superman_sav("superman_answer_key.sav")
export_interpersonal_sav("interpersonal_answer_key.sav")
```

### Subset of Variables (Student Version)

Use `dplyr::select()` to choose only the variables students need, then pipe the result to `export_sav()`. All available labels are automatically preserved for the selected variables:
```{r}
#| eval: false
# Superman — only the variables needed for a specific lab
superman |>
  select(num, media, type, clark_height, rt_critics_score) |>
  export_sav(path = "superman_student.sav")

# Interpersonal survey — just demographics and a few scales
interpersonal_data |>
  select(age, gender, race, gcb, risc, lsas) |>
  export_sav(path = "interpersonal_student.sav")

# Huskers — select with tidyselect helpers
huskers |>
  select(date, season, opp, result, starts_with("ne_")) |>
  export_sav(path = "huskers_nebraska_stats.sav")
```

::: {.callout-tip}
When you pipe a data frame directly to `export_sav()`, it still replaces `NA` with `-99` and declares `-99` as user-missing for all numeric columns. Variable labels and value labels are applied for any columns that have them defined in the package's internal label functions.
:::


## Export All Datasets at Once

To create `.sav` files for every dataset in the package:
```{r}
#| eval: false
export_all_sav(dir = "~/Desktop/PSYC350_SPSS/")
```

This creates one `.sav` file per dataset in the specified folder, using the default filenames. The folder is created if it doesn't exist.


## What Gets Converted During Export

### Categorical Variables

In R, categorical variables are human-readable character strings:
```{r}
superman |>
  select(num, media, type, age_grp) |>
  head(3)
```

In the exported SPSS file, these become numeric codes with value labels:

| Variable | R value | SPSS code | SPSS value label |
|----------|---------|-----------|------------|
| `type` | `"Film"` | 1 | Film |
| `type` | `"TV Series"` | 2 | TV Series |
| `type` | `"Serial"` | 3 | Serial |
| `age_grp` | `"Minimal"` | 1 | Minimal (<2 years) |
| `age_grp` | `"Average"` | 2 | Average (2–5 years) |
| `age_grp` | `"Big"` | 3 | Big (>5 years) |


### Missing Values

In R, missing values are `NA`:
```{r}
superman |>
  select(num, media, ldb_scores) |>
  filter(is.na(ldb_scores))
```

In the exported SPSS file, these become `-99` with the missing value attribute set, so SPSS automatically excludes them from analyses — no manual setup needed.


### Variable Labels

Every variable gets a descriptive label visible in SPSS Variable View. For example:

| Variable | SPSS variable label |
|----------|---------------------|
| `clark_height` | Height of Clark Kent/Superman actor in meters |
| `rt_critics_score` | Rotten Tomatoes critics score (0–100 scale) |
| `age_grp` | Age difference category between Superman and Lois Lane actors |


## Controlling Missing Value Behavior

By default, `NA` values are converted to `-99` with SPSS user-defined missing value codes. This is the standard behavior expected for PSYC 350 labs.

If you instead want SPSS system-missing (`.` in SPSS, which is the equivalent of `NA`), set `use_sentinel = FALSE`:
```{r}
#| eval: false
export_superman_sav("superman_sysmis.sav", use_sentinel = FALSE)
export_sav("superman", path = "superman_sysmis.sav", use_sentinel = FALSE)
```

::: {.callout-note}
System-missing is how SPSS represents truly blank/unknown values. User-defined missing (`-99`) is a specific number that SPSS is told to treat as missing. Both are excluded from analyses, but `-99` is more common in psychology lab settings because it makes it visually obvious that a value is missing when scrolling through Data View.
:::


## Auditing an Export

After exporting, use `check_sav_export()` to verify that labels and missing values are correctly applied:
```{r}
#| eval: false
export_huskers_sav("huskers_data.sav")
check_sav_export("huskers_data.sav")
```

This prints a summary report flagging any issues — for example, variables with missing variable labels or `-99` values that aren't declared as user-missing.

To see details for every variable (not just those with issues):
```{r}
#| eval: false
check_sav_export("huskers_data.sav", show_all = TRUE)
```


## Reading an Exported File Back into R

If you need to read an exported `.sav` file back into R — for example, to verify what students will see — use `get_spss_data()`:
```{r}
#| eval: false
huskers_clean <- get_spss_data("huskers_data.sav")
```

This reads the `.sav` file, converts `-99` back to `NA`, and removes all SPSS labels, giving you a clean tibble ready for R analysis.


## Tips for Instructors

Here's a practical workflow for preparing lab materials:

1. **Generate the answer key** by exporting the full dataset:
```{r}
#| eval: false
export_football_sav("football_answer_key.sav")
```

2. **Generate the student file** by selecting only the variables students need:
```{r}
#| eval: false
football |>
  select(group, volume) |>
  export_sav(path = "football_student.sav")
```

3. **Verify both files** to make sure labels and missing values look correct:
```{r}
#| eval: false
check_sav_export("football_answer_key.sav", show_all = TRUE)
check_sav_export("football_student.sav", show_all = TRUE)
```

4. **Distribute** the student `.sav` file through Canvas or your LMS. Students can open it in SPSS (or JASP) and the value labels, variable labels, and missing value definitions are all pre-configured — no manual data setup required.


## Using Exported Files in SPSS

After opening the `.sav` file in SPSS:

1. **Check Variable View** to see all variable labels and value labels
2. **Toggle "Value Labels"** in the toolbar (or View → Value Labels) to see category names instead of numeric codes in Data View
3. **Missing values are pre-configured** — SPSS will automatically exclude `-99` values from all analyses without any manual setup
4. **Re-export anytime** from R if you need to update the data or change which variables are included
