---
title: "Getting Started with psych350data"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Getting Started with psych350data}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>"
)
```

## Overview

The psych350data package provides all the datasets you need for PSYC 350 labs. Each dataset supports three workflows:

1. **Raw data in R** — Categorical variables are human-readable character strings (like `"Film"`, `"Control"`). Use this for exploration, plotting with ggplot2, and any analysis that doesn't need to match SPSS output.

2. **Prepped for numeric R analysis** — `prep_*()` functions convert character categories to numeric codes that match SPSS exactly. Use this when you need your R output to match SPSS output from a lab assignment.

3. **Exported to SPSS** — `export_*_sav()` functions produce fully labeled `.sav` files with numeric codes, value labels, variable labels, and `-99` for missing values. Use this to create the SPSS file for a lab.

## Loading the Package
```{r setup}
library(psych350data)
library(dplyr)
```

## Browsing Available Datasets
```{r}
list_datasets()
```

For detailed documentation on any dataset:
```{r eval = FALSE}
?superman
?hotones
?mock_jury
?hindsight_mg_data
```

---

## Workflow 1: Raw Data in R

Every dataset is a tibble available immediately after loading the package. Categorical variables use descriptive character values:
```{r}
superman |>
  select(num, media, type, clark_grp, height_gap) |>
  head()
```
```{r}
football |>
  count(group)
```

This format works well with ggplot2 since axis labels are already readable. Missing values are standard `NA`, which R handles automatically:
```{r}
mean(superman$rt_critics_score, na.rm = TRUE)
```

For analyses like ANOVA that need factor inputs, convert directly from the raw data:
```{r}
superman |>
  mutate(
    height_gap = factor(height_gap, levels = c("Minimal", "Average", "Big"))
  ) |>
  pull(height_gap) |>
  levels()
```

---

## Workflow 2: Prep for Numeric R Analysis

If your analysis needs to produce numeric output that matches SPSS — for example, when comparing R results to SPSS output from a lab assignment — use `prep_*()` functions to convert categorical variables to the same numeric codes SPSS uses.

### The `prep_data()` Function

`prep_data()` is the main entry point. Pass it a dataset and the dataset name as a string:
```{r}
superman_prepped <- prep_data(superman, "superman")

superman_prepped |>
  select(num, media, type, clark_grp, height_gap) |>
  head()
```

Each character category is replaced with its numeric code. For example, `type` becomes `1` for Film, `2` for TV Series, and `3` for Serial — the same codes used in the SPSS `.sav` file.

### Dataset-Specific Prep Functions

Every dataset also has its own dedicated prep function:
```{r eval = FALSE}
superman          <- prep_superman(superman)
hotones           <- prep_hotones(hotones)
hotones_sauces    <- prep_hotones_sauces(hotones_sauces)
hotones_episodes  <- prep_hotones_episodes(hotones_episodes)
mcu               <- prep_mcu(mcu)
mock_jury         <- prep_mock_jury(mock_jury)
tip_jokes         <- prep_tip_jokes(tip_jokes)
candy             <- prep_candy(candy)
candy_simple      <- prep_candy_simple(candy_simple)
football          <- prep_football(football)
huskers           <- prep_huskers(huskers)
cheese_data       <- prep_cheese(cheese_data)
lpd_data          <- prep_lpd(lpd_data)
parent_child_data <- prep_parent_child(parent_child_data)
hindsight_mg_data <- prep_hindsight_mg(hindsight_mg_data)
hindsight_wg_data <- prep_hindsight_wg(hindsight_wg_data)
```

### Keeping Labels for Plotting

By default, prep functions replace character columns with numeric codes and the original strings are lost. If you need the original values for axis labels or legends in plots, use `keep_labels = TRUE`. This retains the original character column alongside the numeric one with a `_label` suffix:
```{r}
superman_prepped <- prep_superman(superman, keep_labels = TRUE)

superman_prepped |>
  select(num, type, type_label) |>
  head()
```

This is useful when you want numeric codes for analysis but readable labels for ggplot2:
```{r eval = FALSE}
superman_prepped |>
  ggplot(aes(x = type_label, y = rt_critics_score)) +
  geom_boxplot()
```

---

## Workflow 3: Export to SPSS

Use `export_*_sav()` functions to create `.sav` files for use in SPSS or JASP. The export functions handle everything automatically — you do not need to run a prep function first.

Each export produces a file with:

- Numeric codes for all categorical variables
- Value labels so SPSS displays category names
- Variable labels describing each column
- `-99` for missing values, registered as user-defined missing in SPSS
```{r eval = FALSE}
export_superman_sav("superman_data.sav")
export_football_sav("~/Desktop/football_data.sav")
export_hotones_sav("hotones_data.sav")
export_hindsight_mg_sav("hindsight_mg_data.sav")
export_hindsight_wg_sav("hindsight_wg_data.sav")
```

Or use the generic function with any dataset name:
```{r eval = FALSE}
export_sav("superman", path = "superman_data.sav")
export_sav("hotones")
```

See `vignette("exporting-spss")` for the full export guide including subsetting variables, auditing exports, and controlling missing value behavior.

---

## Choosing the Right Workflow

| Situation | Recommended workflow |
|-----------|----------------------|
| Exploring data, plotting with ggplot2 | Raw data as-is |
| R analysis that needs to match SPSS output | `prep_*()` |
| Plotting after prepping (need readable labels) | `prep_*(..., keep_labels = TRUE)` |
| Creating a `.sav` file for SPSS or JASP | `export_*_sav()` |
| Exporting a subset of variables to SPSS | `select()` then `export_sav()` |

---

## R vs SPSS: What's Different?

| Aspect | Raw R data | After `prep_*()` | SPSS `.sav` |
|--------|-----------|-------------------|-------------|
| Categorical values | Character strings | Numeric codes | Numeric codes with value labels |
| Missing values | `NA` | `NA` | `-99` with user-defined missing |
| Variable descriptions | `?dataset` help files | `?dataset` help files | Variable labels in Variable View |
