---
title: "Getting Started with psych350data"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Getting Started with psych350data}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---
```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The psych350data package provides all the datasets used in PSYC 350 labs at UNL. Every dataset is ready to use the moment you load the package — no file downloading, no manual data entry, and no SPSS license needed.

The package supports **three workflows** depending on what you need to do:

1. **Use the raw R data directly** — categorical variables are human-readable strings like `"Film"` or `"Control"`, which work perfectly with ggplot2 and most R functions.

2. **Prep for numeric analysis matching SPSS** — `prep_*()` functions convert character categories to the same numeric codes used in the SPSS files (e.g., 1, 2, 3), so your R output matches SPSS output exactly. This is useful when working with functions from the [psych350lab](https://github.com/emmarshall/psych350lab) package that expect numeric grouping variables.

3. **Export to SPSS** — `export_*_sav()` functions create fully labeled `.sav` files with numeric codes, value labels, variable labels, and `-99` for missing values. Use this to create answer key files or student data files for labs.


## Installation
```{r}
#| eval: false
# Install from GitHub (one time only)
# install.packages("pak")
pak::pak("emmarshall/psych350data")
```


## Loading the Package
```{r setup}
#| message: false
library(psych350data)
library(dplyr)
```


## Browsing Available Datasets

Use `list_datasets()` to see every dataset in the package:
```{r}
list_datasets()
```

For detailed documentation on any dataset, use `?` in the R console:
```{r}
#| eval: false
?superman
?hotones
?mock_jury
?football
?interpersonal_data
```

See the [Dataset Reference](dataset-reference.qmd) for a complete guide to every dataset including sources and variable descriptions.

---

## Workflow 1: Raw Data in R

Every dataset is a tibble that's available immediately after loading the package. Categorical variables use descriptive character values that are easy to read and plot:
```{r}
superman |>
  select(num, media, type, clark_grp, age_grp) |>
  head()
```
```{r}
football |>
  count(group)
```

### Plotting with ggplot2

Because categorical variables are already human-readable strings, they work directly as axis labels and legend entries in ggplot2 — no extra formatting needed:
```{r}
#| eval: false
library(ggplot2)

# Bar chart — group labels appear automatically
ggplot(football, aes(x = group, y = volume)) +
  geom_boxplot() +
  labs(x = "Group", y = "Brain Volume")

# The labels "Control", "Football no concussion", etc.
# appear on the x-axis without any extra work.
```

If you need to control the **order** of categories on the axis (e.g., for a specific ordering in a bar chart), convert to a factor with explicit levels:
```{r}
#| eval: false
football |>
  mutate(
    group = factor(group, levels = c("Control",
                                     "Football no concussion",
                                     "Football with concussion"))
  ) |>
  ggplot(aes(x = group, y = volume)) +
  geom_boxplot()
```

### Handling Missing Values

Missing values in the raw R data are standard `NA`, which R handles automatically in most functions:
```{r}
# NA values are excluded with na.rm = TRUE
mean(superman$rt_critics_score, na.rm = TRUE)
```

### Running Analyses with Raw Data

For analyses like ANOVA or t-tests that need a factor grouping variable, wrap the character column in `factor()`:
```{r}
# One-way ANOVA using the character group variable
model <- aov(volume ~ factor(group), data = football)
summary(model)
```

---

## Workflow 2: Prep for Numeric Analysis

If your analysis needs to produce output with **numeric codes that match SPSS** — for example, when comparing your R results to SPSS output in a lab, or when using functions from the [psych350lab](https://github.com/emmarshall/psych350lab) package — use `prep_*()` functions.

### What `prep_*()` Does

Each `prep_*()` function replaces character categorical variables with the same numeric codes used in the SPSS `.sav` file. For example, in the `superman` dataset, `type` goes from `"Film"` / `"TV Series"` / `"Serial"` to `1` / `2` / `3`, and `age_grp` goes from `"Minimal"` / `"Average"` / `"Big"` to `1` / `2` / `3`.
```{r}
superman_num <- prep_superman(superman)

superman_num |>
  select(num, media, type, clark_grp, age_grp) |>
  head()
```

### Using the Generic `prep_data()` Function

You can also use the generic `prep_data()` function with any dataset name as a string:
```{r}
superman_num <- prep_data(superman, "superman")

superman_num |>
  select(num, media, type, age_grp) |>
  head()
```

### All Prep Functions

Every dataset has its own dedicated prep function:
```{r}
#| eval: false
prep_superman(superman)
prep_superman_smes(superman_smes)
prep_superman_movies(superman_movies)
prep_hotones(hotones)
prep_hotones_sauces(hotones_sauces)
prep_hotones_episodes(hotones_episodes)
prep_mcu(mcu)
prep_mock_jury(mock_jury)
prep_tip_jokes(tip_jokes)
prep_candy(candy)
prep_candy_simple(candy_simple)
prep_football(football)
prep_huskers(huskers)
prep_interpersonal(interpersonal_data)
prep_self_descriptive(self_descriptive_data)
prep_parent_child(parent_child_data)
prep_hindsight_bg(hindsight_mg_data)
prep_hindsight_wg(hindsight_wg_data)
prep_cheese(cheese_data)
prep_lpd(lpd_data)
```

::: {.callout-note}
Some datasets (like `mock_jury`, `tip_jokes`, `interpersonal_data`, and `self_descriptive_data`) already store their categorical variables as numeric codes. Their prep functions still exist for consistency, but they return the data unchanged.
:::

### Compatibility with psych350lab

The [psych350lab](https://github.com/emmarshall/psych350lab) package provides helper functions for PSYC 350 lab assignments. Many of these functions expect data that looks like an SPSS `.sav` file — that is, categorical/grouping variables stored as **numbers** (not character strings) and missing values coded as `-99` rather than `NA`.

To make any psych350data dataset work with psych350lab functions, you need two steps:

**Step 1: Convert categories to numeric codes** with `prep_*()`:
```{r}
football_num <- prep_football(football)
football_num |> count(group)
# group is now 1, 2, 3 instead of "Control", "Football no concussion", ...
```

**Step 2: Replace NA with -99** (if your psych350lab function expects `-99` for missing):
```{r}
#| eval: false
library(psych350lab)

# Replace NA with -99 across all numeric columns
football_spss <- football_num |>
  mutate(across(where(is.numeric), \(x) ifelse(is.na(x), -99, x)))

# Now the data looks exactly like it would in SPSS
# and is ready for psych350lab functions
```

::: {.callout-tip}
If you're only running standard R functions like `aov()`, `t.test()`, or `cor.test()`, you do **not** need the `-99` step — R handles `NA` natively. The `-99` replacement is only needed for functions that specifically expect SPSS-style missing value coding.
:::

### Keeping Labels for Plotting After Prep

By default, `prep_*()` replaces the character values with numbers, and the original labels are lost. If you need both the numeric codes (for analysis) and the original labels (for plotting), use `keep_labels = TRUE`:
```{r}
superman_both <- prep_superman(superman, keep_labels = TRUE)

superman_both |>
  select(num, type, type_label) |>
  head()
```

This is useful when you want numeric codes for analysis but readable labels in ggplot2:
```{r}
#| eval: false
# Use the _label column for readable axis labels
superman_both |>
  ggplot(aes(x = type_label, y = rt_critics_score)) +
  geom_boxplot() +
  labs(x = "Media Type", y = "Rotten Tomatoes Critics Score")
```

---

## Workflow 3: Export to SPSS

Use `export_*_sav()` functions to create `.sav` files for SPSS or JASP. The export functions handle everything automatically — you do **not** need to run a prep function first.

Each export produces a file with:

- **Numeric codes** for all categorical variables
- **Value labels** so SPSS displays category names (e.g., 1 = "Control")
- **Variable labels** describing each column in SPSS Variable View
- **`-99` for missing values**, registered as user-defined missing so SPSS excludes them automatically

### Quick Export
```{r}
#| eval: false
# Export to current working directory
export_superman_sav()

# Export to a specific location
export_football_sav("~/Desktop/football_data.sav")
```

### Exporting for Answer Keys vs. Student Data

A common workflow is to export the **full dataset** as an instructor answer key, and a **subset of variables** as the student version.

**Full dataset (answer key):**
```{r}
#| eval: false
export_superman_sav("superman_answer_key.sav")
```

**Subset of variables (student version):**
```{r}
#| eval: false
# Use select() to choose only the variables students need,
# then pipe to export_sav()
superman |>
  select(num, media, type, clark_height, rt_critics_score) |>
  export_sav(path = "superman_student.sav")

# Works with tidyselect helpers too
interpersonal_data |>
  select(age, gender, race, gcb, risc, lsas) |>
  export_sav(path = "interpersonal_student.sav")
```

### Export All Datasets at Once
```{r}
#| eval: false
export_all_sav(dir = "~/Desktop/PSYC350_SPSS/")
```

This creates one `.sav` file per dataset in the specified folder.

See the [Exporting to SPSS](exporting-spss.qmd) vignette for the full export guide, including how to audit exports and control missing value behavior.

---

## Choosing the Right Workflow

| Situation | Recommended workflow |
|-----------|----------------------|
| Exploring data, making plots with ggplot2 | Use the raw data as-is |
| Running standard R analyses (t-test, ANOVA, correlation) | Use the raw data with `factor()` for grouping variables |
| R analysis that needs to match SPSS numeric output | `prep_*()` |
| Using psych350lab functions that expect numeric groups and -99 | `prep_*()` then replace NA with -99 |
| Plotting after prepping (need readable axis labels) | `prep_*(..., keep_labels = TRUE)` |
| Creating a `.sav` file for SPSS or JASP | `export_*_sav()` |
| Creating a student data file with fewer variables | `select()` then `export_sav()` |
| Creating instructor answer key `.sav` files | `export_*_sav()` with full dataset |

---

## R vs. SPSS: What's Different?

| Aspect | Raw R data | After `prep_*()` | SPSS `.sav` export |
|--------|-----------|-------------------|---------------------|
| Categorical values | Character strings (`"Film"`, `"Control"`) | Numeric codes (1, 2, 3) | Numeric codes with value labels |
| Missing values | `NA` | `NA` | `-99` with user-defined missing |
| Variable descriptions | `?dataset` help files | `?dataset` help files | Variable labels in Variable View |
| Best for | Plotting, exploration | Matching SPSS output in R | SPSS / JASP labs |
