---
output: github_document
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# psyc350data

<!-- badges: start -->
<!-- badges: end -->

psyc350data provides cleaned, documented datasets for PSYC 350 lab exercises. All datasets are available as ready-to-use R tibbles and can be exported as fully labeled SPSS (.sav) files with variable labels, value labels, and defined missing values.

## Installation

Install from GitHub using [pak](https://pak.r-lib.org/):
```{r eval = FALSE}
# install.packages("pak")
pak::pak("yourusername/psyc350data")
```

Or with devtools:
```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("yourusername/psyc350data")
```

## Available Datasets

| Dataset | Description | N | Source |
|---------|------------|---|--------|
| `superman` | Superman actors across media — heights, RT scores, Letterboxd | 11 | RT, Letterboxd, IMDB |
| `superman_smes` | Simulated SMES ratings by height gap group | 47 | Simulated |
| `hot_ones` | Hot Ones guests — demographics, Scoville ratings, YouTube stats | varies | Hot Ones / YouTube |
| `tip_jokes` | Joke/ad card effect on tipping | varies | Gueguen (2002) |
| `mcu` | MCU Infinity Saga films — box office, RT scores | varies | IMDB |
| `mock_jury` | Attractiveness effects on mock jury sentencing | varies | Plaster (1989) |
| `candy` | Candy rankings with ingredient flags and win % (full) | 85 | FiveThirtyEight |
| `candy_simple` | Candy rankings (simplified — 5 variables) | 85 | FiveThirtyEight |
| `affairs` | Extramarital affairs survey (1969) | varies | Fair (1978) |
| `football` | Football concussions and hippocampus volume | varies | Singh et al. (2014) |

## Quick Start
```{r example}
library(psyc350data)

# See all available datasets
list_datasets()

# Use a dataset directly in R
head(superman)

# Quick summary
str(football)
```

## Exporting to SPSS

All datasets can be exported as .sav files with full SPSS metadata:
```{r eval = FALSE}
# Export a single dataset
export_sav("superman", path = "superman_data.sav")

# Export all datasets to a folder
export_all_sav(dir = "~/Desktop/PSYC350_SPSS/")
```

The exported .sav files include:

- **Variable labels** describing each variable
- **Value labels** for categorical variables (e.g., 1 = "Film", 2 = "TV Series")
- **User-defined missing values** (-99) recognized by SPSS

## Learning More
```{r eval = FALSE}
# Browse the package vignettes
vignette("getting-started", package = "psyc350data")
vignette("exporting-spss", package = "psyc350data")

# View documentation for any dataset
?superman
?hot_ones
?mock_jury
```
